---
title: "UKB - Heteroplasmy and mortality - data analysis"
subtitle: "007 - Supplementary figures - supplementary figures 4 and 5"
author: "Yun Soo Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    pandoc_args: [
    "--number-sections",
    "--number-offset=0"
    ]
fontsize: 11pt
---

# Set up  

## Program description  
* This program reads in the final data set to provide analysis for heteroplasmy count and generates supplementary figures 4 and 5.     
  * Step 1: Read in data 
  * Step 2: Heteroplasmy count, stop mutation, MLC score, and mortality   
  
## Versions of R and packages  
```{r analysis_07 set up, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r analysis_07 load libraries, echo = TRUE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(data.table)
library(rms)
library(RColorBrewer)
library(survival)
library(gtsummary)
library(survminer)
```

## Print out settings  
```{r analysis_07 print settings}
options(tibble.print_max = 100)
options(pillar.sigfig = 5)
```

# Load data  
```{r analysis_07 load data}
ukb <- readRDS("../data_cleaning/ukb_heteroplasmy_mortality_v1.rds")

ukb %<>%
  mutate(
    mutation_nonsyn_any = ifelse(is.na(mutation_nonsyn_any), 0, mutation_nonsyn_any)
  )

table(ukb$smk_na, useNA = "ifany")
table(ukb$het_count, useNA = "ifany")
table(ukb$het_count_cat, useNA = "ifany")

ukb_miss <- ukb %>%
  filter(!is.na(smk_na))

knots <- attributes(rms::rcs(ukb_miss$age, df = 4))$parms
knots
```

# Suvival analysis    
## Settings for survival analysis  
### Time to event  
```{r analysis_07 time_os}
table(ukb$death, useNA = "ifany") %>%
  addmargins()
table(ukb_miss$death, useNA = "ifany") %>%
  addmargins() 

ukb_miss %<>%
  mutate(
    time_os = as.numeric((death_end_fu - dov)/365.25)
  )
summary(ukb_miss$time_os)
```

### Survival object  
```{r analysis_07 os_obj}
os_obj <- with(ukb_miss, Surv(time = time_os, event = death))
```

# Extended figure 4: Heteroplasmy count + stop mutation  
## Figure 4a table    
```{r analysis_07 ext_fig4a_table}
hc_table <- table(ukb_miss$het_count_cat, ukb_miss$death, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(total = alive + died) %>%
  rownames_to_column("Categories") %>%
  add_row() %>%
  mutate(
    row = c(2:6, 1)
  ) %>%
  arrange(row) %>%
  mutate(
    Categories = c("", "0", "1", "2", "3", "4+")
  )
hc_table

stop_table <- table(ukb_miss$mutation_stop_any, ukb_miss$death, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(total = alive + died) %>%
  rownames_to_column("Categories") %>%
  add_row() %>%
  mutate(
    row = c(8, 9, 7)
  ) %>%
  arrange(row) 
stop_table

hc_stop_table <- rbind(hc_table, stop_table)
hc_stop_table %<>%
  mutate(
    Categories = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(Categories)
    ),
    alive = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(alive)
    ),
    died = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(died)
    ),
    total = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(total)
    )
  )
```

## Figure 4b table  
```{r analysis_07 ext_fig4b_table}
mtscore_table <- table(ukb_miss$death, useNA = "ifany") %>%
  as.data.frame() %>%
  spread(Var1, Freq) %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(
    total = alive + died
  ) %>%
  rownames_to_column("Categories") %>%
  add_row() %>%
  mutate(
    row = c(8, 7)
  ) %>%
  arrange(row) 
mtscore_table

hc_mtscore_table <- rbind(hc_table, mtscore_table)
hc_mtscore_table %<>%
  mutate(
    Categories = case_when(
      row == 1 | row == 7 ~ "",
      row == 8 ~ "Continuous", 
      TRUE ~ as.character(Categories)
    ),
    alive = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(alive)
    ),
    died = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(died)
    ),
    total = case_when(
      row == 1 | row == 7 ~ "",
      TRUE ~ as.character(total)
    )
  )
hc_mtscore_table
```

## Cox models  
```{r analysis_07 ext_fig4_cox}
# Figure 4a
os_hc_stop <- os_obj ~ as.factor(het_count_cat) + mutation_stop_any + strata(center) + rcs(age, df= 4) + sex + smk 
hc_stop <- survival::coxph(os_hc_stop, data = ukb_miss)
summary(hc_stop)

coef_hc_stop <- summary(hc_stop)$coef[1:5, 1:3]
coef_hc_stop

coef_hc_stop %<>%
  as_tibble() %>%
  mutate(
    row = c(3:6, 9)
  ) %>%
  add_row(
    row = c(1:2, 7:8)
  ) %>%
  arrange(row)

# Figure 4b
os_hc_mtscore <- os_obj ~ as.factor(het_count_cat) + mito_lc_score + strata(center) + rcs(age, df= 4) + sex + smk
hc_mtscore <- survival::coxph(os_hc_mtscore, data = ukb_miss)
summary(hc_mtscore)

coef_hc_mtscore <- summary(hc_mtscore)$coef[1:5, 1:3]
coef_hc_mtscore

coef_hc_mtscore %<>%
  as_tibble() %>%
  mutate(
    row = c(3:6, 8)
  ) %>%
  add_row(
    row = c(1:2, 7)
  ) %>%
  arrange(row)
```

## Combine result outputs  
### Heteroplasmy count + stop mutation   
```{r analysis_07 combine outputs hc_stop}
hc_stop_data <- hc_stop_table %>%
  left_join(coef_hc_stop, by = "row") %>%
  mutate(
    Variable = 
      c("Heteroplasmy count", "", "", "", "", "", "Nonsense mutation", "", ""),
    lci = 
      ifelse(
        row == 1 | row == 2 | row == 7 | row == 8, 1, exp(coef - 1.96*(`se(coef)`))
      ), 
    uci = 
      ifelse(
        row == 1 | row == 2 | row == 7 | row == 8, 1, exp(coef + 1.96*(`se(coef)`))
      ), 
    ` ` = paste(rep(" ", 25), collapse = " "),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci),
    `HR (95% CI)` = case_when(
      row == 1 | row == 7 ~ "",
      row == 2 | row == 8 ~ "Reference",
      TRUE ~ `HR (95% CI)`
    )
  ) %>%
  rename(
    Total = total,
    Events = died
  ) %>%
  dplyr::select(Variable, Categories, Total, Events, ` `, `HR (95% CI)`, everything())

hc_stop_data
```

### Heteroplasmy count + Mitoscore local  
```{r analysis_07 combine outputs hc_mtscore}
hc_mtscore_data <- hc_mtscore_table %>%
  left_join(coef_hc_mtscore, by = "row") %>%
  mutate(
    Variable = 
      c("Heteroplasmy count", "", "", "", "", "", "MLC score", ""),
    lci = 
      ifelse(
        row == 1 | row == 2 | row == 7, 1, exp(coef - 1.96*(`se(coef)`))
      ), 
    uci = 
      ifelse(
        row == 1 | row == 2 | row == 7, 1, exp(coef + 1.96*(`se(coef)`))
      ), 
    ` ` = paste(rep(" ", 25), collapse = " "),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci),
    `HR (95% CI)` = case_when(
      row == 1 | row == 7 ~ "",
      row == 2 ~ "Reference",
      TRUE ~ `HR (95% CI)`
    )
  ) %>%
  rename(
    Total = total,
    Events = died
  ) %>%
  dplyr::select(Variable, Categories, Total, Events, ` `, `HR (95% CI)`, everything())

hc_mtscore_data
```

## Plot 
### Supplementary Figure 4a  
```{r analysis_07 sfig4a}
sfig4a <- forestploter::forest(
  hc_stop_data[, c(1:6)],
  est = hc_stop_data$`exp(coef)`,
  lower = hc_stop_data$lci,
  upper = hc_stop_data$uci,
  sizes = 0.4,
  ci_column = 5,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.9, 3), 
  ticks_at = c(1, 1.5, 2.0, 2.5)
  )
sfig4a
```

### Supplementary Figure 4b  
```{r analysis_07 sfig4b}
sfig4b <- forestploter::forest(
  hc_mtscore_data[, c(1:6)],
  est = hc_mtscore_data$`exp(coef)`,
  lower = hc_mtscore_data$lci,
  upper = hc_mtscore_data$uci,
  sizes = 0.4,
  ci_column = 5,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.9, 3), 
  ticks_at = c(1, 1.5, 2.0, 2.5)
  )
sfig4b
```

### Supplementary Figure 4  
```{r analysis_07 sfig3, fig.height=7, fig.width=10}
sfig4 <- ggpubr::ggarrange(
  sfig4a, sfig4b,
  ncol = 1, nrow = 2,
  labels = c("a", "b"),
  align = "v"
)
sfig4

ggsave("ukb_heteroplasmy_outcomes_extended_fig4.tiff", plot = sfig4, height=7, width=10) 
```

# Extended figure 5: Heteroplasmy count + nonsynonymous + MLC   
## Figure 5 table    
```{r analysis_07 ext_fig5_table}
hc_table <- table(ukb_miss$het_count_cat, ukb_miss$death, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(total = alive + died) %>%
  rownames_to_column("Categories") %>%
  add_row() %>%
  mutate(
    row = c(2:6, 1)
  ) %>%
  arrange(row) %>%
  mutate(
    Categories = c("", "0", "1", "2", "3", "4+")
  )
hc_table

nonsyn_table <- table(ukb_miss$mutation_nonsyn_any, ukb_miss$death, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(total = alive + died) %>%
  rownames_to_column("Categories") %>%
  add_row() %>%
  mutate(
    row = c(8, 9, 7)
  ) %>%
  arrange(row) 
nonsyn_table

mtscore_table <- table(ukb_miss$death, useNA = "ifany") %>%
  as.data.frame() %>%
  spread(Var1, Freq) %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(
    total = alive + died
  ) %>%
  rownames_to_column("Categories") %>%
  add_row() %>%
  mutate(
    row = c(11, 10)
  ) %>%
  arrange(row) 
mtscore_table

hc_nonsyn_mss_table <- rbind(hc_table, nonsyn_table, mtscore_table)
hc_nonsyn_mss_table %<>%
  mutate(
    Categories = case_when(
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ as.character(Categories)
    ),
    Categories = case_when(
      row == 8 ~ "No", 
      row == 9 ~ "Yes",
      row == 11 ~ "Continuous",
      TRUE ~ as.character(Categories)
    ),
    alive = case_when(
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ as.character(alive)
    ),
    died = case_when(
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ as.character(died)
    ),
    total = case_when(
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ as.character(total)
    )
  )
hc_nonsyn_mss_table
```

## Cox models  
```{r analysis_07 ext_fig5_cox}
os_hc_nonsyn_mtscore <- os_obj ~ as.factor(het_count_cat) + mutation_nonsyn_any + mito_lc_score + strata(center) + rcs(age, df= 4) + sex + smk
hc_nonsyn_mtscore <- survival::coxph(os_hc_nonsyn_mtscore, data = ukb_miss)
summary(hc_nonsyn_mtscore)

coef_hc_nonsyn_mtscore <- summary(hc_nonsyn_mtscore)$coef[1:6, 1:3]
coef_hc_nonsyn_mtscore

coef_hc_nonsyn_mtscore %<>%
  as_tibble() %>%
  mutate(
    row = c(3:6, 9, 11)
  ) %>%
  add_row(
    row = c(1:2, 7:8, 10)
  ) %>%
  arrange(row)
coef_hc_nonsyn_mtscore
```

## Combine result outputs  
```{r analysis_07 combine outputs hc_nonsyn_mss_data}
hc_nonsyn_mss_data <- hc_nonsyn_mss_table %>%
  left_join(coef_hc_nonsyn_mtscore, by = "row") %>%
  mutate(
    Variable = 
      c("Heteroplasmy count", "", "", "", "", "", "Nonsynonymous", "", "", "MLC score", ""),
    lci = 
      ifelse(
        row == 1 | row == 2 | row == 7 | row == 8 | row == 10, 1, exp(coef - 1.96*(`se(coef)`))
      ), 
    uci = 
      ifelse(
        row == 1 | row == 2 | row == 7 | row == 8 | row == 10, 1, exp(coef + 1.96*(`se(coef)`))
      ), 
    ` ` = paste(rep(" ", 25), collapse = " "),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci),
    `HR (95% CI)` = case_when(
      row == 1 | row == 7 | row == 8 | row == 10 ~ "",
      row == 2 | row == 8 ~ "Reference",
      TRUE ~ `HR (95% CI)`
    )
  ) %>%
  rename(
    Total = total,
    Events = died
  ) %>%
  dplyr::select(Variable, Categories, Total, Events, ` `, `HR (95% CI)`, everything())

hc_nonsyn_mss_data
```

## Supplementary Figure 5  
```{r analysis_07 sfig5, fig.height=4, fig.width=8}
sfig5 <- forestploter::forest(
  hc_nonsyn_mss_data[, c(1:6)],
  est = hc_nonsyn_mss_data$`exp(coef)`,
  lower = hc_nonsyn_mss_data$lci,
  upper = hc_nonsyn_mss_data$uci,
  sizes = 0.4,
  ci_column = 5,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.8, 2), 
  ticks_at = c(0.8, 1, 1.5, 2.0)
  )
sfig5

ggsave("ukb_heteroplasmy_outcomes_extended_fig5.tiff", plot = sfig5, height=4, width=8) 
```
