---
title: "UKB - Heteroplasmy and mortality - data analysis"
subtitle: "002 - heteroplasmy count - Figures 3 and 4"
author: "Yun Soo Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    pandoc_args: [
    "--number-sections",
    "--number-offset=0"
    ]
fontsize: 11pt
---

# Set up  

## Program description  
* This program reads in the final data set to provide analysis for heteroplasmy count and generates Figures 3 and 4.     
  * Step 1: Read in data 
  * Step 2: Age, smoking, and heteroplasmy count  
  * Step 3: Heteroplasmy count, stop mutation, MSS, and mortality  
  
## Versions of R and packages  
```{r analysis_02 set up, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r analysis_02 load libraries, echo = TRUE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(data.table)
library(rms)
library(RColorBrewer)
library(survival)
library(gtsummary)
library(survminer)
```

## Print out settings  
```{r analysis_02 print settings}
options(tibble.print_max = 100)
options(pillar.sigfig = 5)
```

# Load data  
```{r analysis_02 load data}
ukb <- readRDS("../data_cleaning/ukb_heteroplasmy_mortality_v1.rds")

table(ukb$smk_na, useNA = "ifany")
table(ukb$het_count, useNA = "ifany")
table(ukb$het_count_cat, useNA = "ifany")

knots <- attributes(rms::rcs(ukb$age, df = 4))$parms
knots

ukb_miss <- ukb %>%
  filter(!is.na(smk_na))
```

# Age, smoking, and heteroplasmy count  
## Regression model    
### Sex  
```{r analysis_02 age sex het_count}
null <- glm(het_count ~ sex, family = "poisson", data = ukb)
summary(null)

age_sex <- glm(het_count ~ rcs(age, knots) + sex + smk_na, family = "poisson", data = ukb)
summary(age_sex)
```

### Smoking  
```{r analysis_02 age smk het_count}
null <- glm(het_count ~ rcs(age, knots) + smk_na, family = "poisson", data = ukb)
summary(null)

age_smk <- glm(het_count ~ rcs(age, knots)*smk_na, family = "poisson", data = ukb)
summary(age_smk)

# p for interaction  
lrtest(null, age_smk)
```

## Figure 3  
```{r analysis_02 figure3}
fig3 <- ukb %>%
  mutate(
    age_cat = case_when(
      age <= 45 ~ 45,
      age > 45 & age <= 50 ~ 50,
      age > 50 & age <= 55 ~ 55,
      age > 55 & age <= 60 ~ 60,
      age > 60 & age <= 65 ~ 65,
      TRUE ~ 75
    )
  ) %>%
  filter(!is.na(smk_na)) %>%
  ggplot() +
  geom_point(aes(x = factor(age_cat), y = het_count, color = smk),
             stat = "summary", fun = "mean", size = 3) +
  geom_errorbar(aes(x = factor(age_cat), y = het_count, color = smk),
                stat = "summary", fun.data = "mean_se", fun.args = list(mult = 1.96), width = 0) +
  scale_x_discrete("Age category (years)",
                   labels = c(expression("" <= 45), "46-50", "51-55", "55-60", "61-65", "> 65")) +
  scale_y_continuous("Mean heteroplasmy count \n (95% CI)") +
  scale_color_manual(
    values = c("navyblue", "orange", "red"),
    labels = c("Never smoker", "Former smoker", "Current smoker")
    ) +
  theme_classic() +
  theme(legend.title = element_blank(),
        text = element_text(size = 15)) 
fig3

ggsave("ukb_heteroplasmy_outcomes_fig3.tiff", plot = fig3)
```

# Heteroplasmy count and mortality  
## Settings for survival analysis  
### Time to event  
```{r analysis_02 time_os}
table(ukb$death, useNA = "ifany") %>%
  addmargins()
table(ukb_miss$death, useNA = "ifany") %>%
  addmargins() 

ukb %<>%
  mutate(
    time_os = as.numeric((death_end_fu - dov)/365.25)
  )
summary(ukb$time_os)
```

### Survival object  
```{r analysis_02 os_obj}
os_obj <- with(ukb, Surv(time = time_os, event = death))
```

## Cox models  
### Heteroplasmy count  
```{r analysis_02 os_hc}
hc_table <- table(ukb_miss$het_count_cat, ukb_miss$death, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rename(
    alive = `0`,
    died = `1` 
  ) %>%
  mutate(
    total = alive + died
  )
hc_table 

os_hc <- os_obj ~ as.factor(het_count_cat) + strata(center) + rcs(age, df= 4) + sex + smk  
hc <- survival::coxph(os_hc, data = ukb)
summary(hc)

coef_hc <- summary(hc)$coef[1:4, 1:3]
coef_hc
```

### Stop mutation  
```{r analysis_02 os_stop}
stop_table <- table(ukb_miss$mutation_stop_any, ukb_miss$death, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rename(
    alive = `0`,
    died = `1` 
  ) %>%
  mutate(
    total = alive + died
  )
stop_table 

os_stop <- os_obj ~ mutation_stop_any + strata(center) + rcs(age, df= 4) + sex + smk 
stop <- survival::coxph(os_stop, data = ukb)
summary(stop)

coef_stop <- summary(stop)$coef[1, 1:3] 
coef_stop
```

### MLC score sum   
```{r analysis_02 os_mss}
mss_table <- table(ukb_miss$death, useNA = "ifany") %>%
  as.data.frame() %>%
  spread(Var1, Freq) %>%
  rename(
    alive = `0`,
    died = `1`
  ) %>%
  mutate(
    total = alive + died
  )
mss_table

os_mss <- os_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk
mss <- survival::coxph(os_mss, data = ukb)
summary(mss)

coef_mss <- summary(mss)$coef[1, 1:3]
coef_mss
```

### Heteroplasmy count, stop mutation, and MSS  
```{r analysis_02 os_het_stop_mss}
os_het_stop_mss <- os_obj ~ as.factor(het_count_cat) + mutation_stop_any + mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk
het_stop_mss <- survival::coxph(os_het_stop_mss, data = ukb)
summary(het_stop_mss)

coef_adj <- summary(het_stop_mss)$coef[1:6, 1:3]
coef_adj
```

## Combine result outputs  
```{r analysis_02 combine outputs}
# combine unadjusted estimates to a single table  
coef_unadj <- rbind(coef_hc, coef_stop, coef_mss) %>%
  as_tibble() %>%
  mutate(
    var = c("hc1", "hc2", "hc3", "hc4", "stop", "mss")
  ) %>%
  rename(
    coef_unadj = coef,
    exp_coef_unadj = `exp(coef)`,
    se_coef_unadj = `se(coef)`
  )
coef_unadj
  
# to tibble and data transformation  
coef_adj %<>%
  as_tibble() %>%
  mutate(
    var = c("hc1", "hc2", "hc3", "hc4", "stop", "mss")
  ) %>%
  rename(
    coef_adj = coef,
    exp_coef_adj = `exp(coef)`,
    se_coef_adj = `se(coef)`
  )
coef_adj

# merge the adjusted and unadjusted estimates  
coef_all <- coef_unadj %>%
  left_join(coef_adj, by = "var") %>%
  # start with `var` variable  
  select(var, everything()) %>%
  # add 2 additional rows for categorical variables (1 for group name, 1 for reference) and 1 additional row for continuous variable (1 for group name)
  mutate(
    row = c(3:6, 9, 11)
  ) %>%
  add_row(
    row = c(1:2, 7:8, 10)
  ) %>%
  arrange(row) %>%
  # add a column for group and category  
  mutate(
    Variable = 
      c("Heteroplasmy count", "", "", "", "", "", "Nonsense mutation", "", "", "MLC score sum (MSS)", ""),
    Categories = 
      c("", "0", "1", "2", "3", "4+", "", "No", "Yes", "", "Continuous")
  ) %>%
  # tidy the unadjusted estimates  
  mutate(
    lci_unadj =
      ifelse(row == 1 | row == 2 | row == 7 | row == 8 | row == 10, 1, exp(coef_unadj - 1.96*se_coef_unadj)),
    uci_unadj =
      ifelse(row == 1 | row == 2 | row == 7 | row == 8 | row == 10, 1, exp(coef_unadj + 1.96*se_coef_unadj)),
    `Unadjusted` = paste(rep(" ", 25), collapse = " "),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", exp_coef_unadj, lci_unadj, uci_unadj),
    `HR (95% CI)` = case_when(
      row == 2 | row == 8 ~ "Reference",
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ `HR (95% CI)`)
    ) %>%
  # tidy the adjusted estimates  
  mutate(
    lci_adj = 
      ifelse(row == 1 | row == 2 | row == 7 | row == 8 | row == 10, 1, exp(coef_adj - 1.96*se_coef_adj)),
    uci_adj = 
      ifelse(row == 1 | row == 2 | row == 7 | row == 8 | row == 10, 1, exp(coef_adj + 1.96*se_coef_adj)),
    `Mutually adjusted` = paste(rep(" ", 25), collapse = " "),
    `Adjusted HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", exp_coef_adj, lci_adj, uci_adj),
    `Adjusted HR (95% CI)` = case_when(
      row == 2 | row == 8 ~ "Reference",
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ `Adjusted HR (95% CI)`)
    ) %>%
  dplyr::select(
    Variable, Categories, exp_coef_unadj, lci_unadj, uci_unadj, `Unadjusted`, `HR (95% CI)`, exp_coef_adj, lci_adj, uci_adj, `Mutually adjusted`, `Adjusted HR (95% CI)`, row
  )

coef_all 
``` 

## Combine count tables  
```{r analysis_02 combine count}
table <- rbind(hc_table, stop_table, mss_table) %>%
  as_tibble() %>%
  mutate(
    row = c(2:6, 8:9, 11)
  ) %>%
  add_row(
    row = c(1, 7, 10)
  ) %>%
  arrange(row)
table
```

## Merge results table with count table 
```{r analysis_02 merge tables}
data <- coef_all %>%
  left_join(table, by = "row") %>%
  mutate(
    total = as.character(total), 
    total = case_when(
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ total
    ),
    died = as.character(died),
    died = case_when(
      row == 1 | row == 7 | row == 10 ~ "",
      TRUE ~ died
    ) 
  ) %>%
  rename(
    Total = total,
    Events = died
  ) %>%
  dplyr::select(
    Variable, Categories, Total, Events, `Unadjusted`, `HR (95% CI)`, `Mutually adjusted`, `Adjusted HR (95% CI)`, exp_coef_unadj, lci_unadj, uci_unadj, exp_coef_adj, lci_adj, uci_adj 
  )
data
```

# Figure 4  
```{r analysis_02 figure4, fig.height=4.5, fig.width=15}
fig4 <- forestploter::forest(
  data[, c(1:8)],
  est = list(
    data$exp_coef_unadj,
    data$exp_coef_adj
    ),
  lower = list(
    data$lci_unadj,
    data$lci_adj
    ),
  upper = list(
    data$uci_unadj,
    data$uci_adj
    ),
  sizes = 0.6,
  ci_column = c(5, 7),
  ref_line = c(1, 1),
  xlog = TRUE,
  xlim = list(
    c(0.8, 3), 
    c(0.8, 3)
    ),
  ticks_at = list(
    c(0.8, 1, 1.5, 2.0, 2.5, 3),
    c(0.8, 1, 1.5, 2.0, 2.5, 3)
    )
  )
fig4  

ggsave("ukb_heteroplasmy_outcomes_fig4.tiff", plot = fig4, height=4.5, width=13)
```


