---
title: "UKB - Heteroplasmy and mortality - data analysis"
subtitle: "004 - Mito score sum and cancer - Figure 7"
author: "Yun Soo Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    pandoc_args: [
    "--number-sections",
    "--number-offset=0"
    ]
fontsize: 11pt
---

# Set up  

## Program description  
* This program reads in the final data set to provide analysis for mitoscore sum (MSS) and cancer and generates Figure 7.     
  * Step 1: Read in data 
  * Step 2: Prevalent cancer     
  * Step 3: Incident cancer  
  * Step 4: Mortality in prevalent or incident cancer  
  
## Versions of R and packages  
```{r analysis_04 set up, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r analysis_04 load libraries, echo = TRUE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(data.table)
library(rms)
library(RColorBrewer)
library(survival)
library(gtsummary)
library(survminer)
library(RStata)
```

## Print out settings  
```{r analysis_04 print settings}
options(tibble.print_max = 100)
options(pillar.sigfig = 5)

# chooseStataBin()
options("RStata.StataPath" = "\"C:\\Program Files\\Stata16\\StataMP-64\"")
options("RStata.StataVersion" = 16)
```

# Load data  
```{r analysis_04 load data}
ukb <- readRDS("../data_cleaning/ukb_heteroplasmy_mortality_v1.rds")
```

# Case definitions for prevalent disease  
```{r analysis_04 case prev}
ukb %<>%
  mutate(
    prev_cancer_resp = ifelse(cancer_dx_resp == 1 & cancer_date_resp < dov, 1, 0),
    prev_cancer_breast = ifelse(cancer_dx_breast == 1 & cancer_date_breast < dov, 1, 0), 
    prev_cancer_hemato_lymphoma = ifelse(cancer_dx_hemato_lymphoma == 1 & cancer_date_hemato_lymphoma < dov, 1, 0),
    prev_cancer_hemato_leukemia = ifelse(cancer_dx_hemato_leukemia == 1 & cancer_date_hemato_leukemia < dov, 1, 0)
    ) 

ukb %<>%
  mutate(
    time_prev_resp = as.numeric((cancer_date_resp - dov)/365.25),
    time_prev_breast = as.numeric((cancer_date_breast - dov)/365.25),
    time_prev_hemato_lymphoma = as.numeric((cancer_date_hemato_lymphoma - dov)/365.25),
    time_prev_hemato_leukemia = as.numeric((cancer_date_hemato_leukemia - dov)/365.25)
  )
```

# Case definitions for incident disease  
```{r analysis_04 case inc}
ukb %<>%
  mutate(
    inc_cancer_resp = ifelse(cancer_dx_resp == 1 & cancer_date_resp >= dov, 1, 0),
    inc_cancer_breast = ifelse(cancer_dx_breast == 1 & cancer_date_breast >= dov, 1, 0), 
    inc_cancer_hemato_lymphoma = ifelse(cancer_dx_hemato_lymphoma == 1 & cancer_date_hemato_lymphoma >= dov, 1, 0),
    inc_cancer_hemato_leukemia = ifelse(cancer_dx_hemato_leukemia == 1 & cancer_date_hemato_leukemia >= dov, 1, 0)
  ) 

ukb %<>%
  mutate(
    time_inc_resp = as.numeric((cancer_date_resp - dov)/365.25),
    time_inc_breast = as.numeric((cancer_date_breast - dov)/365.25),
    time_inc_hemato_lymphoma = as.numeric((cancer_date_hemato_lymphoma - dov)/365.25),
    time_inc_hemato_leukemia = as.numeric((cancer_date_hemato_leukemia - dov)/365.25)
  ) 
``` 

# Indicator variable for prevalent/incident cancer  
```{r analysis_04 prev inc}
ukb %<>%
  mutate(
    prev_inc_resp = 
      case_when(
        prev_cancer_resp == 1 ~ 0,
        inc_cancer_resp == 1 ~ 1
      ),
    prev_inc_breast = 
      case_when(
        prev_cancer_breast == 1 ~ 0, 
        inc_cancer_breast == 1 ~ 1
      ),
    prev_inc_hemato_lymphoma = 
      case_when(
        prev_cancer_hemato_lymphoma == 1 ~ 0,
        inc_cancer_hemato_lymphoma == 1 ~ 1
      ),
    prev_inc_hemato_leukemia = 
      case_when(
        prev_cancer_hemato_leukemia == 1 ~ 0,
        inc_cancer_hemato_leukemia == 1 ~ 1
      )
  ) 
```

# Prevalent cancer  
## Table of events  
```{r analysis_04 prev_table}
ukb_prev <- ukb %>%
  filter(!is.na(smk_na) & !is.na(alc_na) & !is.na(bmi) & !is.na(wbc) & !is.na(hap))

prev_table <- 
  apply(ukb_prev[c("prev_cancer_resp", "prev_cancer_breast", "prev_cancer_hemato_lymphoma", "prev_cancer_hemato_leukemia")], 2, table) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Cancer type") %>%
  mutate(
    `Cancer type` = case_when(
      `Cancer type` == "prev_cancer_resp" ~ "Lung", 
      `Cancer type` == "prev_cancer_breast" ~ "Breast",
      `Cancer type` == "prev_cancer_hemato_lymphoma" ~ "Lymphoma", 
      `Cancer type` == "prev_cancer_hemato_leukemia" ~ "Leukemia"
    )
  ) %>%
  rename(Events = `1`) %>%
  select(`Cancer type`, Events)
prev_table
```

## By type of cancer  
### Malignant neoplasms of respiratory and intrathoracic organs  
```{r analysis_04 prev_lung}
model <- prev_cancer_resp ~ mito_score_sum_lc + center + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
lr_cancer_resp <- glm(model, 
                      family = binomial, 
                      data = ukb_prev) 

# marginal prevalence 
mp_cancer_resp <- summary(prediction::prediction(lr_cancer_resp, 
                                                 at = list(mito_score_sum_lc = c(0, 1))))
mp_cancer_resp

# prevalence ratio
pr_cancer_resp <- mp_cancer_resp$Prediction[2]/mp_cancer_resp$Prediction[1]
pr_cancer_resp

# 95% confidence interval using delta method 
var_cancer_resp_ref <- (mp_cancer_resp$SE[1]^2)/(mp_cancer_resp$Prediction[1]^2)
var_cancer_resp_comp <- (mp_cancer_resp$SE[2]^2)/(mp_cancer_resp$Prediction[2]^2)
se_logpr_mito_score_sum_lc <- sqrt(var_cancer_resp_ref + var_cancer_resp_comp)

pr_cancer_resp_lci <- exp(log(pr_cancer_resp) - 1.96*se_logpr_mito_score_sum_lc)
pr_cancer_resp_uci <- exp(log(pr_cancer_resp) + 1.96*se_logpr_mito_score_sum_lc)

prev_lung <- c(pr_cancer_resp, pr_cancer_resp_lci, pr_cancer_resp_uci)
prev_lung
```

### Malignant neoplasm of breast  
```{r analysis_04 prev_breast}
model <- prev_cancer_breast ~ mito_score_sum_lc + center + rcs(age, df= 4) + smk + alc + bmi + wbc + hap
lr_cancer_breast <- glm(model, 
                        family = binomial, 
                        data = ukb_prev,
                        subset = sex == "women") 

# marginal prevalence 
mp_cancer_breast <- summary(prediction::prediction(lr_cancer_breast,
                                                   at = list(mito_score_sum_lc = c(0, 1))))
mp_cancer_breast

# prevalence ratio
pr_cancer_breast <- mp_cancer_breast$Prediction[2]/mp_cancer_breast$Prediction[1]
pr_cancer_breast

# 95% confidence interval using delta method 
var_cancer_breast_ref <- (mp_cancer_breast$SE[1]^2)/(mp_cancer_breast$Prediction[1]^2)
var_cancer_breast_comp <- (mp_cancer_breast$SE[2]^2)/(mp_cancer_breast$Prediction[2]^2)
se_logpr_mito_score_sum_lc <- sqrt(var_cancer_breast_ref + var_cancer_breast_comp)

pr_cancer_breast_lci <- exp(log(pr_cancer_breast) - 1.96*se_logpr_mito_score_sum_lc)
pr_cancer_breast_uci <- exp(log(pr_cancer_breast) + 1.96*se_logpr_mito_score_sum_lc)

prev_breast <- c(pr_cancer_breast, pr_cancer_breast_lci, pr_cancer_breast_uci)
prev_breast
```

### Lymphomas    
```{r analysis_04 prev_lymphoma}
model <- prev_cancer_hemato_lymphoma ~ mito_score_sum_lc + center + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap
lr_cancer_hemato_lymphoma <- glm(model, 
                                 family = binomial,
                                 data = ukb_prev) 

# marginal prevalence 
mp_cancer_hemato_lymphoma <- summary(prediction::prediction(lr_cancer_hemato_lymphoma,
                                                            at = list(mito_score_sum_lc = c(0, 1))))
mp_cancer_hemato_lymphoma

# prevalence ratio
pr_cancer_hemato_lymphoma <- mp_cancer_hemato_lymphoma$Prediction[2]/mp_cancer_hemato_lymphoma$Prediction[1]
pr_cancer_hemato_lymphoma

# 95% confidence interval using delta method 
var_cancer_hemato_lymphoma_ref <- (mp_cancer_hemato_lymphoma$SE[1]^2)/(mp_cancer_hemato_lymphoma$Prediction[1]^2)
var_cancer_hemato_lymphoma_comp <- (mp_cancer_hemato_lymphoma$SE[2]^2)/(mp_cancer_hemato_lymphoma$Prediction[2]^2)
se_logpr_mito_score_sum_lc <- sqrt(var_cancer_hemato_lymphoma_ref + var_cancer_hemato_lymphoma_comp)

pr_cancer_hemato_lymphoma_lci <- exp(log(pr_cancer_hemato_lymphoma) - 1.96*se_logpr_mito_score_sum_lc)
pr_cancer_hemato_lymphoma_uci <- exp(log(pr_cancer_hemato_lymphoma) + 1.96*se_logpr_mito_score_sum_lc)

prev_lymphoma <- c(pr_cancer_hemato_lymphoma, pr_cancer_hemato_lymphoma_lci, pr_cancer_hemato_lymphoma_uci)
prev_lymphoma
```

### Leukemias  
```{r analysis_04 prev_leukemia}
model <- prev_cancer_hemato_leukemia ~ mito_score_sum_lc + center + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap
lr_cancer_hemato_leukemia <- glm(model,
                                 family = binomial,
                                 data = ukb_prev) 

# marginal prevalence 
mp_cancer_hemato_leukemia <- summary(prediction::prediction(lr_cancer_hemato_leukemia, 
                                                            at = list(mito_score_sum_lc = c(0, 1))))
mp_cancer_hemato_leukemia

# prevalence ratio
pr_cancer_hemato_leukemia <- mp_cancer_hemato_leukemia$Prediction[2]/mp_cancer_hemato_leukemia$Prediction[1]
pr_cancer_hemato_leukemia

# 95% confidence interval using delta method 
var_cancer_hemato_leukemia_ref <- (mp_cancer_hemato_leukemia$SE[1]^2)/(mp_cancer_hemato_leukemia$Prediction[1]^2)
var_cancer_hemato_leukemia_comp <- (mp_cancer_hemato_leukemia$SE[2]^2)/(mp_cancer_hemato_leukemia$Prediction[2]^2)
se_logpr_mito_score_sum_lc <- sqrt(var_cancer_hemato_leukemia_ref + var_cancer_hemato_leukemia_comp)

pr_cancer_hemato_leukemia_lci <- exp(log(pr_cancer_hemato_leukemia) - 1.96*se_logpr_mito_score_sum_lc)
pr_cancer_hemato_leukemia_uci <- exp(log(pr_cancer_hemato_leukemia) + 1.96*se_logpr_mito_score_sum_lc)

prev_leukemia <- c(pr_cancer_hemato_leukemia, pr_cancer_hemato_leukemia_lci, pr_cancer_hemato_leukemia_uci)
prev_leukemia
```

## Combine results table  
```{r analysis_04 prev results}
prev_results <- rbind(prev_lung, prev_breast, prev_lymphoma, prev_leukemia) %>%
  as.data.frame() %>%
  rename(
    coef = V1,
    lci = V2, 
    uci = V3
  ) 
prev_results
```

## Merge table and results  
```{r analysis_04 prev table results}
data_prev <- bind_cols(
  prev_table, prev_results
  ) %>%
  mutate(
    `PR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", coef, lci, uci),
    ` ` = paste(rep(" ", 25), collapse= " ") 
  ) %>%
  dplyr::select(
    `Cancer type`, Events, ` `, `PR (95% CI)`, everything()
  ) 

data_prev
```

## Figure 7a  
```{r analysis_04 figure7a, fig.height=3, fig.width=7}
fig7a <- forestploter::forest(
  data_prev[, c(1:4)],
  est = data_prev$coef,
  lower = data_prev$lci,
  upper = data_prev$uci,
  sizes = 0.4,
  ci_column = 3,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.7, 6), 
  ticks_at = c(0.75, 1, 1.5, 2.0, 5),
  title = "Prevalence ratios for each cancer type"
  )
fig7a
```

# Incident cancer  
## Case definition  
* Combine incident and mortality as an outcome definition  
```{r analysis_04 inc case definition}
ukb %<>%
  mutate(
    yn_cancer = 
      case_when(
        cancer_dx_cancer == 1 | cod_cancer == 1 ~ 1, 
        TRUE ~ 0
      ),
    yn_resp = 
      case_when(
        cancer_dx_resp == 1 | cod_cancer_resp == 1 ~ 1,
        TRUE ~ 0
      ),
    yn_breast =
      case_when(
        cancer_dx_breast == 1 | cod_cancer_breast == 1 ~ 1,
        TRUE ~ 0
      ),
    yn_hemato_lymphoma = 
      case_when(
        cancer_dx_hemato_lymphoma == 1 | cod_cancer_lymphoma == 1 ~ 1,
        TRUE ~ 0
      ),
    yn_hemato_leukemia = 
      case_when(
        cancer_dx_hemato_leukemia == 1 | cod_cancer_leukemia == 1 ~ 1,
        TRUE ~ 0
      )
  )
```

## Dates of each event  
* The date of event for each case is defined as the minimum of the two dates (`cancer_date_*` and `date_of_death`)  
* When the two are equal, we keep the `cancer_date_*`  
* For those who died without having a neoplasm, the date of event is `date_of_death`  
* For those who did not have neoplasm and did not die, the date of event is the date of administrative censoring  
```{r analysis_04 inc dates}
adm_censor <- as.Date("2021-11-12")

ukb %<>%
  mutate(
    date_cancer = 
      as.Date(
        ifelse(yn_cancer == 1 & ((cancer_date_cancer <= date_of_death) | death == 0),
               cancer_date_cancer, 
               ifelse(yn_cancer == 0 & death == 1, date_of_death, adm_censor)), 
        origin = "1970-01-01"
      ),
    date_resp = 
      as.Date(
        ifelse(yn_resp == 1 & ((cancer_date_resp <= date_of_death) | death == 0),
               cancer_date_resp,
               ifelse(yn_resp == 0 & death == 1, date_of_death, adm_censor)),
        origin = "1970-01-01"
      ),
    date_breast = 
      as.Date(
        ifelse(yn_breast == 1 & ((cancer_date_breast <= date_of_death) | death == 0),
               cancer_date_breast,
               ifelse(yn_breast == 0 & death == 1, date_of_death, adm_censor)),
        origin = "1970-01-01"
      ),
    date_hemato_lymphoma = 
      as.Date(
        ifelse(yn_hemato_lymphoma == 1 & ((cancer_date_hemato_lymphoma <= date_of_death) | death == 0),
               cancer_date_hemato_lymphoma,
               ifelse(yn_hemato_lymphoma == 0 & death == 1, date_of_death, adm_censor)),
        origin = "1970-01-01"
      ),
    date_hemato_leukemia = 
      as.Date(
        ifelse(yn_hemato_leukemia == 1 & ((cancer_date_hemato_leukemia <= date_of_death) | death == 0),
               cancer_date_hemato_leukemia,
               ifelse(yn_hemato_leukemia == 0 & death == 1, date_of_death, adm_censor)),
        origin = "1970-01-01"
      )
    ) 
```

## Exclusion of prevalent cases  
* Exclude from the analysis if an individual had **any** cancer prior to assessment visit  (n = 13,701)  
* In addition, also exclude 1272 participants when the date of diagnosis is missing for any of the outcomes.  
```{r analysis_04 inc exclude prev}
ukb %>%
  count(date_cancer < dov)

ukb_inc <- ukb %>% ## (n = 180528)
  filter(date_cancer >= dov &
          !is.na(date_cancer) & !is.na(date_resp) & !is.na(date_breast) & !is.na(date_hemato_lymphoma) & !is.na(date_hemato_leukemia)
         ) %>%
  filter(!is.na(smk_na) & !is.na(alc_na) & !is.na(bmi) & !is.na(wbc) & !is.na(hap))
```

## Table of events  
```{r analysis_04 inc_table}
inc_table <- 
  apply(ukb_inc[c("yn_resp", "yn_breast", "yn_hemato_lymphoma", "yn_hemato_leukemia")], 2, table) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Cancer type") %>%
  mutate(
    `Cancer type` = case_when(
      `Cancer type` == "yn_resp" ~ "Lung", 
      `Cancer type` == "yn_breast" ~ "Breast",
      `Cancer type` == "yn_hemato_lymphoma" ~ "Lymphoma", 
      `Cancer type` == "yn_hemato_leukemia" ~ "Leukemia"
    )
  ) %>%
  rename(Events = `1`) %>%
  select(`Cancer type`, Events)
inc_table
```

## Settings for survival analysis   
### Time to event    
```{r analysis_04 inc time to event}
ukb_inc %<>%
  mutate(
    time_resp = as.numeric((date_resp - dov)/365.25),
    time_breast = as.numeric((date_breast - dov)/365.25),
    time_hemato_lymphoma = as.numeric((date_hemato_lymphoma - dov)/365.25),
    time_hemato_leukemia = as.numeric((date_hemato_leukemia - dov)/365.25)
  ) 
```

### Survival object  
### Malignant neoplasms of respiratory and intrathoracic organs  
```{r analysis_04 cs_resp_obj}
cs_resp_obj <- with(ukb_inc, Surv(time = time_resp, event = yn_resp))

model <- cs_resp_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap
inc_lung <- survival::coxph(model, data = ukb_inc) 
summary(inc_lung)

coef_lung <- summary(inc_lung)$coef[1, 1:3]
coef_lung
``` 

### Malignant neoplasm of breast  
```{r analysis_04 cs_breast_obj}
cs_breast_obj <- with(ukb_inc, Surv(time = time_breast, event = yn_breast))

model <- cs_breast_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + smk + alc + bmi + wbc + hap
inc_breast <- survival::coxph(model, data = ukb_inc, subset = sex == "women") 
summary(inc_breast)

coef_breast <- summary(inc_breast)$coef[1, 1:3]
coef_breast
```

#### Lymphomas  
```{r analysis_04 cs_hemato_lymphoma_obj}
cs_hemato_lymphoma_obj <- with(ukb_inc, Surv(time = time_hemato_lymphoma, event = yn_hemato_lymphoma))

model <- cs_hemato_lymphoma_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap
inc_lymphoma <- survival::coxph(model, data = ukb_inc) 
summary(inc_lymphoma)

coef_lymphoma <- summary(inc_lymphoma)$coef[1, 1:3]
coef_lymphoma
```

#### Leukemias  
```{r analysis_04 cs_hemato_leukemia_obj}
cs_hemato_leukemia_obj <- with(ukb_inc, Surv(time = time_hemato_leukemia, event = yn_hemato_leukemia))

model <- cs_hemato_leukemia_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap
inc_leukemia <- survival::coxph(model, data = ukb_inc) 
summary(inc_leukemia)

coef_leukemia <- summary(inc_leukemia)$coef[1, 1:3]
coef_leukemia
```

## Combine results table  
```{r analysis_04 inc results}
inc_results <- rbind(coef_lung, coef_breast, coef_lymphoma, coef_leukemia) %>%
  as.data.frame() %>%
  mutate(
    lci = exp(coef - 1.96*`se(coef)`),
    uci = exp(coef + 1.96*`se(coef)`),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci),
    ` ` = paste(rep(" ", 25), collapse= " ") 
  )
inc_results
```

## Merge table and results  
```{r analysis_04 inc table results}
data_inc <- bind_cols(
  inc_table, inc_results
  ) %>%
  dplyr::select(
    `Cancer type`, Events, ` `, `HR (95% CI)`, everything()
  ) 

data_inc
```

## Figure 7b  
```{r analysis_04 figure7b, fig.height=3, fig.width=7}
fig7b <- forestploter::forest(
  data_inc[, c(1:4)],
  est = data_inc$`exp(coef)`,
  lower = data_inc$lci,
  upper = data_inc$uci,
  sizes = 0.4,
  ci_column = 3,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.7, 6), 
  ticks_at = c(0.75, 1, 1.5, 2.0, 5),
  title = "Hazard ratios for each cancer type"
  )
fig7b
```

# Mortality in cancer patients  
* This model uses Fine and Gray method for competing risks. As late entries are not allowed in R, we use Stata to run the models.  

## Case definitions for competing event    
* event of interest is coded as `1`, death due to (other) cancers is coded as `2`, and non-cancer deaths (including benign neoplasms) are coded as `3`.  

### Malignant neoplasms of respiratory and intrathoracic organs  
```{r analysis_04 death event_resp}
ukb_death <- ukb %>% 
  filter(!is.na(smk_na) & !is.na(alc_na) & !is.na(bmi) & !is.na(wbc) & !is.na(hap))

ukb_death %<>%
  mutate(
    event_resp = 
      case_when(
        death == 0 ~ 0, 
        cod_cancer_resp == 1 ~ 1,
        cod_cancer == 1 & cod_cancer_resp == 0 ~ 2,
        cod_infection == 1 | (cod_neoplasm == 1 & cod_cancer == 0) | cod_blood == 1 | cod_endocrine == 1 | cod_mental == 1 |
          cod_neuro == 1 | cod_ear == 1 | cod_cardio == 1 | cod_pulmo == 1 | cod_gi == 1 | cod_skin == 1 | cod_muscle == 1 | 
          cod_gu == 1 | cod_pregnancy == 1 | cod_congenital == 1 | cod_uns == 1 | cod_special == 1 | cod_external == 1 ~ 3
      )
  ) 

table_lung <- table(ukb_death$prev_inc_resp, ukb_death$event_resp) %>%
  unclass() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("type") %>%
  mutate(
    Any = `0` + `1`
  ) %>%
  rename(
    Prevalent = `0`,
    Incident = `1`
  )  %>%
  filter(type == 1) %>%
  add_row(
    Any = table(ukb_death$prev_inc_resp)[1] + table(ukb_death$prev_inc_resp)[2],
    Prevalent = table(ukb_death$prev_inc_resp)[1],
    Incident = table(ukb_death$prev_inc_resp)[2]
  ) %>%
  dplyr::select(-type) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Prevalent or incident") %>%
  mutate(
    "Cancer type" = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Lung"
  ) %>%
  rename(
    Events = V1,
    Total = V2
  ) %>%
  mutate(
    order = c(3, 4, 2, 1),
    `Prevalent or incident` = case_when(
      `Cancer type` == "Lung" ~ "",
      TRUE ~ `Prevalent or incident`
    ),
    Total = case_when(
      `Cancer type` == "Lung" ~ "",
      TRUE ~ as.character(Total)
    ),
    Events = case_when(
      `Cancer type` == "Lung" ~ "",
      TRUE ~ as.character(Events)
    )
  ) %>%
  arrange(order) %>%
  dplyr::select(`Cancer type`, `Prevalent or incident`, Total, Events)
table_lung
```

### Malignant neoplasm of breast    
```{r analysis_04 death event_breast}
ukb_death %<>%
  mutate(
    event_breast = 
      case_when(
        death == 0 ~ 0, 
        cod_cancer_breast == 1 ~ 1,
        cod_cancer == 1 & cod_cancer_breast == 0 ~ 2,
        cod_infection == 1 | (cod_neoplasm == 1 & cod_cancer == 0) | cod_blood == 1 | cod_endocrine == 1 | cod_mental == 1 |
          cod_neuro == 1 | cod_ear == 1 | cod_cardio == 1 | cod_pulmo == 1 | cod_gi == 1 | cod_skin == 1 | cod_muscle == 1 | 
          cod_gu == 1 | cod_pregnancy == 1 | cod_congenital == 1 | cod_uns == 1 | cod_special == 1 | cod_external == 1 ~ 3
      )
  ) 


table_breast <- table(ukb_death$prev_inc_breast, ukb_death$event_breast) %>%
  unclass() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("type") %>%
  mutate(
    Any = `0` + `1`
  ) %>%
  rename(
    Prevalent = `0`,
    Incident = `1`
  )  %>%
  filter(type == 1) %>%
  add_row(
    Any = table(ukb_death$prev_inc_breast)[1] + table(ukb_death$prev_inc_breast)[2],
    Prevalent = table(ukb_death$prev_inc_breast)[1],
    Incident = table(ukb_death$prev_inc_breast)[2]
  ) %>%
  dplyr::select(-type) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Prevalent or incident") %>%
  mutate(
    "Cancer type" = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Breast"
  ) %>%
  rename(
    Events = V1,
    Total = V2
  ) %>%
  mutate(
    order = c(3, 4, 2, 1),
    `Prevalent or incident` = case_when(
      `Cancer type` == "Breast" ~ "",
      TRUE ~ `Prevalent or incident`
    ),
    Total = case_when(
      `Cancer type` == "Breast" ~ "",
      TRUE ~ as.character(Total)
    ),
    Events = case_when(
      `Cancer type` == "Breast" ~ "",
      TRUE ~ as.character(Events)
    )
  ) %>%
  arrange(order) %>%
  dplyr::select(`Cancer type`, `Prevalent or incident`, Total, Events)
table_breast
```

### Lymphoma  
```{r analysis_04 death event_lymphoma}
ukb_death %<>%
  mutate(
    event_lymphoma = 
      case_when(
        death == 0 ~ 0, 
        cod_cancer_lymphoma == 1 ~ 1,
        cod_cancer == 1 & cod_cancer_lymphoma == 0 ~ 2,
        cod_infection == 1 | (cod_neoplasm == 1 & cod_cancer == 0) | cod_blood == 1 | cod_endocrine == 1 | cod_mental == 1 |
          cod_neuro == 1 | cod_ear == 1 | cod_cardio == 1 | cod_pulmo == 1 | cod_gi == 1 | cod_skin == 1 | cod_muscle == 1 | 
          cod_gu == 1 | cod_pregnancy == 1 | cod_congenital == 1 | cod_uns == 1 | cod_special == 1 | cod_external == 1 ~ 3
      )
  ) 
table_lymphoma <- table(ukb_death$prev_inc_hemato_lymphoma, ukb_death$event_lymphoma) %>%
  unclass() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("type") %>%
  mutate(
    Any = `0` + `1`
  ) %>%
  rename(
    Prevalent = `0`,
    Incident = `1`
  ) %>%
  filter(type == 1) %>%
  add_row(
    Any = table(ukb_death$prev_inc_hemato_lymphoma)[1] + table(ukb_death$prev_inc_hemato_lymphoma)[2],
    Prevalent = table(ukb_death$prev_inc_hemato_lymphoma)[1],
    Incident = table(ukb_death$prev_inc_hemato_lymphoma)[2]
  ) %>%
  dplyr::select(-type) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Prevalent or incident") %>%
  mutate(
    "Cancer type" = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Lymphoma"
  ) %>%
  rename(
    Events = V1,
    Total = V2
  ) %>%
  mutate(
    order = c(3, 4, 2, 1),
    `Prevalent or incident` = case_when(
      `Cancer type` == "Lymphoma" ~ "",
      TRUE ~ `Prevalent or incident`
    ),
    Total = case_when(
      `Cancer type` == "Lymphoma" ~ "",
      TRUE ~ as.character(Total)
    ),
    Events = case_when(
      `Cancer type` == "Lymphoma" ~ "",
      TRUE ~ as.character(Events)
    )
  ) %>%
  arrange(order) %>%
  dplyr::select(`Cancer type`, `Prevalent or incident`, Total, Events)
table_lymphoma
```

### Leukemia  
```{r analysis_04 death event_leukemia}
ukb_death %<>%
  mutate(
    event_leukemia = 
      case_when(
        death == 0 ~ 0, 
        cod_cancer_leukemia == 1 ~ 1,
        cod_cancer == 1 & cod_cancer_leukemia == 0 ~ 2,
        cod_infection == 1 | (cod_neoplasm == 1 & cod_cancer == 0) | cod_blood == 1 | cod_endocrine == 1 | cod_mental == 1 |
          cod_neuro == 1 | cod_ear == 1 | cod_cardio == 1 | cod_pulmo == 1 | cod_gi == 1 | cod_skin == 1 | cod_muscle == 1 | 
          cod_gu == 1 | cod_pregnancy == 1 | cod_congenital == 1 | cod_uns == 1 | cod_special == 1 | cod_external == 1 ~ 3
      )
  ) 
table_leukemia <- table(ukb_death$prev_inc_hemato_leukemia, ukb_death$event_leukemia) %>%
  unclass() %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("type") %>%
  mutate(
    Any = `0` + `1`
  ) %>%
  rename(
    Prevalent = `0`,
    Incident = `1`
  ) %>%
  filter(type == 1) %>%
  add_row(
    Any = table(ukb_death$prev_inc_hemato_leukemia)[1] + table(ukb_death$prev_inc_hemato_leukemia)[2],
    Prevalent = table(ukb_death$prev_inc_hemato_leukemia)[1],
    Incident = table(ukb_death$prev_inc_hemato_leukemia)[2]
  ) %>%
  dplyr::select(-type) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Prevalent or incident") %>%
  mutate(
    "Cancer type" = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Leukemia"
  ) %>%
  rename(
    Events = V1,
    Total = V2
  ) %>%
  mutate(
    order = c(3, 4, 2, 1),
    `Prevalent or incident` = case_when(
      `Cancer type` == "Leukemia" ~ "",
      TRUE ~ `Prevalent or incident`
    ),
    Total = case_when(
      `Cancer type` == "Leukemia" ~ "",
      TRUE ~ as.character(Total)
    ),
    Events = case_when(
      `Cancer type` == "Leukemia" ~ "",
      TRUE ~ as.character(Events)
    )
  ) %>%
  arrange(order) %>%
  dplyr::select(`Cancer type`, `Prevalent or incident`, Total, Events)
table_leukemia
```

### Combine event table  
```{r analysis_04 death table}
death_table <- rbind(table_lung, table_breast, table_lymphoma, table_leukemia)
death_table
```

## Settings for survival analysis   
### Time to event  
```{r analysis_04 death time to event}
ukb_death %<>%
  mutate(
    entry_resp = as.Date(ifelse(prev_cancer_resp == 1, dov, cancer_date_resp), origin = "1970-01-01"),
    exit_resp = death_end_fu,
    origin_resp = cancer_date_resp,
    time_resp = as.numeric((exit_resp - entry_resp)/365.25)
  )

ukb_death %<>%
  mutate(
    entry_breast = as.Date(ifelse(prev_cancer_breast == 1, dov, cancer_date_breast), origin = "1970-01-01"),
    exit_breast = death_end_fu,
    origin_breast = cancer_date_breast,
    time_breast = as.numeric((exit_breast - entry_breast)/365.25)
  )

ukb_death %<>%
  mutate(
    entry_hemato_lymphoma = as.Date(ifelse(prev_cancer_hemato_lymphoma == 1, dov, cancer_date_hemato_lymphoma), origin = "1970-01-01"),
    exit_hemato_lymphoma = death_end_fu,
    origin_hemato_lymphoma = cancer_date_hemato_lymphoma,
    time_hemato_lymphoma = as.numeric((exit_hemato_lymphoma - entry_hemato_lymphoma)/365.25)
  )

ukb_death %<>%
  mutate(
    entry_hemato_leukemia = as.Date(ifelse(prev_cancer_hemato_leukemia == 1, dov, cancer_date_hemato_leukemia), origin = "1970-01-01"),
    exit_hemato_leukemia = death_end_fu,
    origin_hemato_leukemia = cancer_date_hemato_leukemia,
    time_hemato_leukemia = as.numeric((exit_hemato_leukemia - entry_hemato_leukemia)/365.25)
  )
```

### Survival analysis in Stata
```{r analysis_04 death stata}
stata("heteroplasmy_mortality_analysis_04_mss_cancer.do")
```

### Read in Stata results  
```{r analysis_04 death stata output}
death_results <- readxl::read_excel("heteroplasmy_mortality_analysis_04_mss_cancer_stata_output.xlsx") %>%
  dplyr::select(-`...1`) %>%
  filter(`...2` == "mito_score_sum_lc") 
death_results

# Lung cancer
death_lung <- death_results %>%
  dplyr::select("any_lung...3", "any_lung...4", "prev_lung...11", "prev_lung...12", "inc_lung...19", "inc_lung...20") %>%
  data.table::setnames(
    c("any_lung...3", "any_lung...4", "prev_lung...11", "prev_lung...12", "inc_lung...19", "inc_lung...20"),
    c("any_coef", "any_se", "prev_coef", "prev_se", "inc_coef", "inc_se")
  )  

death_lung %<>%
  pivot_longer(
    everything(),
    names_to = c("Prevalent or incident", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  mutate(
    `Cancer type` = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Lung"
  ) %>%
  mutate(
    `Prevalent or incident` = case_when(
      `Prevalent or incident` == "any" ~ "Any",
      `Prevalent or incident` == "prev" ~ "Prevalent",
      `Prevalent or incident` == "inc" ~ "Incident", 
      TRUE ~ "" 
    ),
    order = c(2, 3, 4, 1)
  ) %>% 
  arrange(order) %>%
  dplyr::select(
    `Cancer type`, `Prevalent or incident`, coef, se 
  )
death_lung

# Breast cancer
death_breast <- death_results %>%
  dplyr::select("any_breast...5", "any_breast...6", "prev_breast...13", "prev_breast...14", "inc_breast...21", "inc_breast...22") %>%
  data.table::setnames(
    c("any_breast...5", "any_breast...6", "prev_breast...13", "prev_breast...14", "inc_breast...21", "inc_breast...22"),
    c("any_coef", "any_se", "prev_coef", "prev_se", "inc_coef", "inc_se")
  )

death_breast %<>%
  pivot_longer(
    everything(),
    names_to = c("Prevalent or incident", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  mutate(
    `Cancer type` = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Breast"
  ) %>%
  mutate(
    `Prevalent or incident` = case_when(
      `Prevalent or incident` == "any" ~ "Any",
      `Prevalent or incident` == "prev" ~ "Prevalent",
      `Prevalent or incident` == "inc" ~ "Incident", 
      TRUE ~ "" 
    ),
    order = c(2, 3, 4, 1)
  ) %>% 
  arrange(order) %>%
  dplyr::select(
    `Cancer type`, `Prevalent or incident`, coef, se 
  )
death_breast

# Lymphoma  
death_lymphoma <- death_results %>%
  dplyr::select("any_lymphoma...7", "any_lymphoma...8", "prev_lymphoma...15", "prev_lymphoma...16", "inc_lymphoma...23", "inc_lymphoma...24") %>%
  data.table::setnames(
    c("any_lymphoma...7", "any_lymphoma...8", "prev_lymphoma...15", "prev_lymphoma...16", "inc_lymphoma...23", "inc_lymphoma...24"),
    c("any_coef", "any_se", "prev_coef", "prev_se", "inc_coef", "inc_se")
  ) 

death_lymphoma %<>%
  pivot_longer(
    everything(),
    names_to = c("Prevalent or incident", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  mutate(
    `Cancer type` = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Lymphoma"
  ) %>%
  mutate(
    `Prevalent or incident` = case_when(
      `Prevalent or incident` == "any" ~ "Any",
      `Prevalent or incident` == "prev" ~ "Prevalent",
      `Prevalent or incident` == "inc" ~ "Incident", 
      TRUE ~ "" 
    ),
    order = c(2, 3, 4, 1)
  ) %>% 
  arrange(order) %>%
  dplyr::select(
    `Cancer type`, `Prevalent or incident`, coef, se 
  )
death_lymphoma

# Leukemia
death_leukemia <- death_results %>%
  dplyr::select("any_leukemia...9", "any_leukemia...10", "prev_leukemia...17", "prev_leukemia...18", "inc_leukemia...25", "inc_leukemia...26") %>%
  data.table::setnames(
    c("any_leukemia...9", "any_leukemia...10", "prev_leukemia...17", "prev_leukemia...18", "inc_leukemia...25", "inc_leukemia...26"),
    c("any_coef", "any_se", "prev_coef", "prev_se", "inc_coef", "inc_se")
  ) 

death_leukemia %<>%
  pivot_longer(
    everything(),
    names_to = c("Prevalent or incident", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  mutate(
    `Cancer type` = c("", "", "")
  ) %>%
  add_row(
    `Cancer type` = "Leukemia"
  ) %>%
  mutate(
    `Prevalent or incident` = case_when(
      `Prevalent or incident` == "any" ~ "Any",
      `Prevalent or incident` == "prev" ~ "Prevalent",
      `Prevalent or incident` == "inc" ~ "Incident", 
      TRUE ~ "" 
    ),
    order = c(2, 3, 4, 1)
  ) %>% 
  arrange(order) %>%
  dplyr::select(
    `Cancer type`, `Prevalent or incident`, coef, se 
  )
death_leukemia
```

## Combine results table  
```{r analysis_04 death results}
death_results <- bind_rows(death_lung, death_breast, death_lymphoma, death_leukemia)
death_results %<>%
  mutate(
    coef = as.numeric(coef),
    se = as.numeric(se),
    `exp(coef)` = exp(coef),
    lci = exp(coef - 1.96*se),
    uci = exp(coef + 1.96*se),
    ` ` = paste(rep(" ", 25), collapse= " "),
    coef = case_when(
      `Cancer type` == "Lung" | `Cancer type` == "Breast" | `Cancer type` == "Lymphoma" | `Cancer type` == "Leukemia" ~ "",
      TRUE ~ as.character(coef)
    ),
    se = case_when(
      `Cancer type` == "Lung" | `Cancer type` == "Breast" | `Cancer type` == "Lymphoma" | `Cancer type` == "Leukemia" ~ "",
      TRUE ~ as.character(se)
    ),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci), 
    `HR (95% CI)` = case_when(
      `Cancer type` == "Lung" | `Cancer type` == "Breast" | `Cancer type` == "Lymphoma" | `Cancer type` == "Leukemia" ~ "",
      TRUE ~ as.character(`HR (95% CI)`)
    )
  ) %>%
  dplyr::select(-c(`Cancer type`, `Prevalent or incident`))
death_results
```

## Merge table and results  
```{r analysis_04 death table results}
data_death <- bind_cols(
  death_table, death_results
  ) %>%
  dplyr::select(
    `Cancer type`, `Prevalent or incident`, `Total`, Events, ` `, `HR (95% CI)`, everything()
  ) 
data_death
```

## Figure 7c  
```{r analysis_04 figure7c, fig.height=6, fig.width=10}
fig7c <- forestploter::forest(
  data_death[, c(1:6)],
  est = data_death$`exp(coef)`,
  lower = data_death$lci,
  upper = data_death$uci,
  sizes = 0.4,
  ci_column = 5,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.7, 6), 
  ticks_at = c(0.75, 1.5, 2, 5),
  title = "Hazard ratios for mortality due to cancer type"
  )
fig7c
```

# Figure 7  
```{r analysis_04 figure7, fig.height=6, fig.width=18}
fig7 <- ggpubr::ggarrange(
  ggarrange(fig7a, fig7b, ncol = 1, labels = c("a", "b")), 
  fig7c,
  ncol = 2, 
  labels = c("", "c"),
  widths = c(0.8, 1)
)
fig7

ggsave("ukb_heteroplasmy_outcomes_fig7.tiff", plot = fig7, height=6, width=16) 
``` 
