---
title: "UKB - Heteroplasmy and mortality - data analysis"
subtitle: "003 - Mito score sum and mortality - Figures 5 and 6"
author: "Yun Soo Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    pandoc_args: [
    "--number-sections",
    "--number-offset=0"
    ]
fontsize: 11pt
---

# Set up  

## Program description  
* This program reads in the final data set to provide analysis for heteroplasmy count and generates Figures 5 and 6.     
  * Step 1: Read in data 
  * Step 2: Dose-response relationship between MSS and mortality   
  * Step 3: MSS and cause of death    
  
## Versions of R and packages  
```{r analysis_03 set up, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r analysis_03 load libraries, echo = TRUE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(data.table)
library(rms)
library(RColorBrewer)
library(survival)
library(gtsummary)
library(survminer)
```

## Print out settings  
```{r analysis_03 print settings}
options(tibble.print_max = 100)
options(pillar.sigfig = 5)
```

# Load data  
```{r analysis_03 load data}
ukb <- readRDS("../data_cleaning/ukb_heteroplasmy_mortality_v1.rds")

ukb_os <- ukb %>%
  filter(!is.na(smk_na) & !is.na(alc_na) & !is.na(bmi) & !is.na(wbc) & !is.na(hap))

knots <- attributes(rms::rcs(ukb_os$mito_score_sum_lc, 4))$parms
knots
```

# Dose-response relationship between MSS and mortality  
## Time to event  
```{r analysis_03 time_os}
ukb_os %<>%
  mutate(
    time_os = as.numeric((death_end_fu - dov)/365.25)
  )
```

## Survival object  
```{r analysis_03 figure5 os_obj}
os_obj <- with(ukb_os, Surv(time = time_os, event = death))
```

## Figure 5
### All-cause mortality  
```{r analysis_03 os_models}
table(ukb_os$death)

model <- os_obj ~ rcs(mito_score_sum_lc, knots) + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap

os_model <- survival::coxph(model, data = ukb_os)
summary(os_model)
matrcs <- model.matrix(os_model)[, 1:3]

beta_rcs <- coef(os_model)[1:3]
beta_rcs

var_rcs <- vcov(os_model)[c(1:3), c(1:3)]
var_rcs

loghr_rcs <- as.vector(matrcs %*% beta_rcs)
loghrse_rcs <- sqrt(diag(matrcs %*% var_rcs %*% t(matrcs)))
loghr_rcs_ll <- as.vector(loghr_rcs - 1.96*loghrse_rcs)
loghr_rcs_ul <- as.vector(loghr_rcs + 1.96*loghrse_rcs)

hrrcsplot <- ukb_os %>%
  dplyr::select(mito_score_sum_lc) %>%
  cbind(loghr_rcs, loghrse_rcs, loghr_rcs_ll, loghr_rcs_ul) 

hrrcsplot %<>%
  mutate(upper = exp(loghr_rcs_ul),
         lower = exp(loghr_rcs_ll),
         hr = exp(loghr_rcs))

hr <- hrrcsplot %>% 
  ggplot() +
  geom_ribbon(aes(x = mito_score_sum_lc, ymin = lower, ymax = upper), fill = "#005cb2", alpha = 0.3) +
  geom_line(aes(x = mito_score_sum_lc, y = hr), linetype = "solid", size = 1.5, color = "#005cb2") +
  geom_hline(yintercept = 1,
             linetype = "dashed",
             size = 1) +
  scale_x_continuous(name = "",
                     expand = c(0.01, 0.01),
                     breaks = seq(0, 2, by = 0.25), 
                     limits = c(0, 2)) +
  scale_y_continuous(name = "Hazard ratio for all-cause mortality",
                     breaks = c(0.9, 1, 1.25, 1.5, 1.75, 2, 2.5),
                     limits = c(0.9, 2.5), 
                     trans = "log") + 
  theme_classic() +
  theme(axis.title = element_text(face = "bold",
                                  size = 14),
        axis.text = element_text(size = 12))
```

### Histogram of MLC score sum (MSS)  
```{r analysis_03 hist_mss}
ukb %>% 
  count(mito_score_sum_lc > 2 & !is.na(mito_score_sum_lc))

hist <- ukb %>%
  filter(het_count != 0 & mito_score_sum_lc <= 2) %>%
  ggplot() +
  geom_histogram(aes(x = mito_score_sum_lc), binwidth = 0.05, fill = "#005cb2") +
  geom_rug(aes(x = mito_score_sum_lc, y = NULL),
           length = unit(0.05, "npc"), 
           color = "black") +
  scale_x_continuous("MLC score sum (MSS)", 
                     expand = c(0.01, 0.01),
                     breaks = seq(0, 2, by = 0.25),
                     limits = c(0, 2)) +
  scale_y_continuous("Count",
                     limits = c(0, 10000),
                     expand = c(0.08, 0)) +
  theme_classic() +
  theme(axis.title = element_text(face = "bold",
                                  size = 14),
        axis.text = element_text(size = 12))
hist
``` 

### Combine plots    
```{r analysis_03, fig.width=8, fig.height=6}
fig5 <- ggpubr::ggarrange(
  hr, hist,
  ncol = 1, nrow = 2,
  heights = c(3, 1),
  align = "hv"
)
ggsave("ukb_heteroplasmy_outcomes_fig5.tiff", plot = fig5) 
```

# MSS and cause of death   
## Cause of death  
### Table for number of events  
```{r analysis_03 cod_table}
cod_table <- table(ukb_os$cod, useNA = "ifany") %>%
  unclass() %>%
  as.data.frame() %>%
  rownames_to_column("Cause of death") %>%
  rename(Events = ".") %>%
  mutate(
    `Cause of death` = case_when(
      `Cause of death` == "alive" ~ "All-cause mortality",
      `Cause of death` == "infection" ~ "Infection",
      `Cause of death` == "neoplasm" ~ "Neoplasm",
      `Cause of death` == "blood" ~ "Benign diseases of the blood", 
      `Cause of death` == "endocrine" ~ "Endocrine disorders", 
      `Cause of death` == "mental" ~ "Mental and behavioral disorders", 
      `Cause of death` == "neuro" ~ "Neurological disorders", 
      `Cause of death` == "cardio" ~ "Circulatory disorders", 
      `Cause of death` == "pulmo" ~ "Respiratory disorders", 
      `Cause of death` == "gi" ~ "Digestive disorders", 
      `Cause of death` == "gu" ~ "Genitourinary disorders", 
      `Cause of death` == "covid" ~ "COVID-19", 
      `Cause of death` == "external" ~ "External causes"
    ),
    Events = ifelse(`Cause of death` == "All-cause mortality", table(ukb_os$death)[2], Events)
  ) %>%
  mutate(
    row = c(1:13)
  )
cod_table
```

### Survival object  
#### All-cause mortality  
```{r analysis_03 os_obj}
os_obj <- with(ukb_os, Surv(time = time_os, event = death))
os_all <- os_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
all <- survival::coxph(os_all, data = ukb_os)
summary(all)
summary(all)$coef

coef_all <- summary(all)$coef[1, ]
coef_all
```

#### Infection  
```{r analysis_03 cs_infection_obj}
cs_infection_obj <- with(ukb_os, Surv(time = time_os, event = cod_infection))
os_infection <- cs_infection_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
infection <- survival::coxph(os_infection, data = ukb_os)
summary(infection)

coef_infection <- summary(infection)$coef[1, ]
coef_infection
```

#### Neoplasm  
```{r analysis_03 cs_neoplasm_obj}
cs_neoplasm_obj <- with(ukb_os, Surv(time = time_os, event = cod_neoplasm))
os_neoplasm <- cs_neoplasm_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
neoplasm <- survival::coxph(os_neoplasm, data = ukb_os)
summary(neoplasm)

coef_neoplasm <- summary(neoplasm)$coef[1, ]
coef_neoplasm
```

#### Benign diseases of the blood    
```{r analysis_03 cs_blood_obj}
cs_blood_obj <- with(ukb_os, Surv(time = time_os, event = cod_blood))
os_blood <- cs_blood_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
blood <- survival::coxph(os_blood, data = ukb_os)
summary(blood)

coef_blood <- summary(blood)$coef[1, ]
coef_blood
```

#### Endocrine disorders  
```{r analysis_03 cs_endocrine_obj}
cs_endocrine_obj <- with(ukb_os, Surv(time = time_os, event = cod_endocrine))
os_endocrine <- cs_endocrine_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
endocrine <- survival::coxph(os_endocrine, data = ukb_os)
summary(endocrine)

coef_endocrine <- summary(endocrine)$coef[1, ]
coef_endocrine
```

#### Mental and behavioral disorders  
```{r analysis_03 cs_mental_obj}
cs_mental_obj <- with(ukb_os, Surv(time = time_os, event = cod_mental))
os_mental <- cs_mental_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
mental <- survival::coxph(os_mental, data = ukb_os)
summary(mental)

coef_mental <- summary(mental)$coef[1, ]
coef_mental
```

#### Neurological disorders  
```{r analysis_03 cs_neuro_obj}
cs_neuro_obj <- with(ukb_os, Surv(time = time_os, event = cod_neuro))
os_neuro <- cs_neuro_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
neuro <- survival::coxph(os_neuro, data = ukb_os)
summary(neuro)

coef_neuro <- summary(neuro)$coef[1, ]
coef_neuro
```

#### Circulatory disorders  
```{r analysis_03 cs_cardio_obj}
cs_cardio_obj <- with(ukb_os, Surv(time = time_os, event = cod_cardio))
os_cardio <- cs_cardio_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cardio <- survival::coxph(os_cardio, data = ukb_os)
summary(cardio)

coef_cardio <- summary(cardio)$coef[1, ]
coef_cardio
```

#### Respiratory disorders  
```{r analysis_03 cs_pulmo_obj}
cs_pulmo_obj <- with(ukb_os, Surv(time = time_os, event = cod_pulmo))
os_pulmo <- cs_pulmo_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
pulmo <- survival::coxph(os_pulmo, data = ukb_os)
summary(pulmo)

coef_pulmo <- summary(pulmo)$coef[1, ]
coef_pulmo
```

#### Digestive disorders  
```{r analysis_03 cs_gi_obj}
cs_gi_obj <- with(ukb_os, Surv(time = time_os, event = cod_gi))
os_gi <- cs_gi_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
gi <- survival::coxph(os_gi, data = ukb_os)
summary(gi)

coef_gi <- summary(gi)$coef[1, ]
coef_gi
```

#### Genitourinary disorders  
```{r analysis_03 cs_gu_obj}
cs_gu_obj <- with(ukb_os, Surv(time = time_os, event = cod_gu))
os_gu <- cs_gu_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
gu <- survival::coxph(os_gu, data = ukb_os)
summary(gu)

coef_gu <- summary(gu)$coef[1, ]
coef_gu
```

#### COVID-19  
```{r analysis_03 cs_covid_obj}
cs_covid_obj <- with(ukb_os, Surv(time = time_os, event = cod_covid))
os_covid <- cs_covid_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
covid <- survival::coxph(os_covid, data = ukb_os)
summary(covid)

coef_covid <- summary(covid)$coef[1, ]
coef_covid
```

#### External causes  
```{r analysis_03 cs_external_obj}
cs_external_obj <- with(ukb_os, Surv(time = time_os, event = cod_external))
os_external <- cs_external_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
external <- survival::coxph(os_external, data = ukb_os)
summary(external)

coef_external <- summary(external)$coef[1, ]
coef_external
```

### Combine results   
```{r analysis_03 cod combine results}
# coefficients 
cod_coef <- rbind(
  coef_all, coef_infection, coef_neoplasm, coef_blood, coef_endocrine, coef_mental, coef_neuro, coef_cardio, coef_pulmo, coef_gi, coef_gu, coef_covid, coef_external
  ) %>%
  as_tibble() %>%
  rename(
    p = `Pr(>|z|)`
  ) 
cod_coef

# calculate FDR 
cod_coef %<>%
  mutate(
    fdr = p.adjust(p, "BH"),
    fdr = ifelse(fdr < 0.05, format(fdr, format = "e", digits = 3), round(fdr, 2)),
    # fdr = round(p.adjust(p, "BH"), 2),
    fdr = as.character(fdr)
  ) %>%
  mutate(
    row = c(1:13)
  )
cod_coef
```

### Merge table with results  
```{r analysis_03 cod merge table results}
data <- cod_table %>%
  left_join(cod_coef, by = "row") %>%
  mutate(
    # add estimates  
    lci = exp(coef - 1.96*`se(coef)`),
    uci = exp(coef + 1.96*`se(coef)`),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci),
    ` ` = paste(rep(" ", 25), collapse= " "), 
    # reformat p-values 
    pp = ifelse(p < 0.05, format(p, format = "e", digits = 3), round(p, 2)),
    pp = as.character(pp),
    pp = case_when(
      pp == "1.30e-12" ~ "1.30x10\u207b\u00b9\u00b2",
      pp == "3.66e-04" ~ "3.66x10\u207b\u2074",
      pp == "2.41e-02" ~ "2.41x10\u207b\u00b2",
      pp == "3.28e-11" ~ "3.28x10\u207b\u00b9\u00b9", 
      pp == "0.2" ~ "0.20",
      pp == "0.3" ~ "0.30",
      TRUE ~ pp
    )
  ) %>%
  rename(
    `P value` = pp
  ) %>%
  dplyr::select(
    `Cause of death`, Events, ` `, `HR (95% CI)`, `P value`, everything()
  ) %>%
  filter(Events >= 100) %>%
  arrange(-coef) %>%
  mutate(
    order = row_number(),
    order = ifelse(
      `Cause of death` == "All-cause mortality", 0, order
    ),
    Events = as.character(Events)
  ) %>%
  arrange(order)
data
```

### Figure 6a  
```{r analysis_03 figure6a, fig.height=4.5, fig.width=9}
fig6a <- forestploter::forest(
  data[, c(1:5)],
  est = data$`exp(coef)`,
  lower = data$lci,
  upper = data$uci,
  sizes = 0.4,
  ci_column = 3,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.5, 3), 
  ticks_at = c(0.5, 0.75, 1, 1.5, 2.0, 3),
  title = "Hazard ratios for each cause of death"
  )
fig6a
```

## By cancer type   
### Table for number of events  
```{r analysis_03 cod_cancer_table}
cod_cancer_table <- 
  apply(ukb_os[c("cod_cancer", "cod_cancer_blood", "cod_cancer_head_neck", "cod_cancer_gi", "cod_cancer_resp",
                 "cod_cancer_bone", "cod_cancer_skin", "cod_cancer_soft_tissue", "cod_cancer_breast",
                 "cod_cancer_female", "cod_cancer_male", "cod_cancer_urinary", "cod_cancer_cns", 
                 "cod_cancer_endo", "cod_cancer_uns", "cod_cancer_lymphoma", "cod_cancer_leukemia")], 2, table) %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("Cancer type") %>%
  mutate(
    `Cancer type` = case_when(
      `Cancer type` == "cod_cancer" ~ "Any cancer",
      `Cancer type` == "cod_cancer_blood" ~ "Hematologic", 
      `Cancer type` == "cod_cancer_head_neck" ~ "Head and neck",
      `Cancer type` == "cod_cancer_gi" ~ "Digestive organs", 
      `Cancer type` == "cod_cancer_resp" ~ "Lung", 
      `Cancer type` == "cod_cancer_bone" ~ "Bone", 
      `Cancer type` == "cod_cancer_skin" ~ "Skin", 
      `Cancer type` == "cod_cancer_soft_tissue" ~ "Mesothelial and soft tissue",
      `Cancer type` == "cod_cancer_breast" ~ "Breast",
      `Cancer type` == "cod_cancer_female" ~ "Female genital organs", 
      `Cancer type` == "cod_cancer_male" ~ "Male genital organs", 
      `Cancer type` == "cod_cancer_urinary" ~ "Urinary tract", 
      `Cancer type` == "cod_cancer_cns" ~ "Eye, brain, and other parts of CNS", 
      `Cancer type` == "cod_cancer_endo" ~ "Endocrine organs",
      `Cancer type` == "cod_cancer_uns" ~ "Ill-defined, secondary and unspecified sites", 
      `Cancer type` == "cod_cancer_lymphoma" ~ "  Lymphoma", 
      `Cancer type` == "cod_cancer_leukemia" ~ "  Leukemia"
    ),
    row = c(1:17)
  ) %>%
  rename(Events = `1`) %>%
  select(`Cancer type`, Events, row)
cod_cancer_table
```

#### Cancer      
```{r analysis_03 cs_cancer_obj}
cs_cancer_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer))
os_cancer <- cs_cancer_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer <- survival::coxph(os_cancer, data = ukb_os)
summary(cancer)

coef_cancer <- summary(cancer)$coef[1, ]
coef_cancer
```

#### Malignant neoplasms of lip, oral cavity, and pharynx  
```{r analysis_03 cs_cancer_head_neck_obj}
cs_cancer_head_neck_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_head_neck))
os_cancer_head_neck <- cs_cancer_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_head_neck <- survival::coxph(os_cancer_head_neck, data = ukb_os)
summary(cancer_head_neck)

coef_cancer_head_neck <- summary(cancer_head_neck)$coef[1, ]
coef_cancer_head_neck
```

#### Malignant neoplasms of digestive organs  
```{r analysis_03 cs_cancer_gi_obj}
cs_cancer_gi_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_gi))
os_cancer_gi <- cs_cancer_gi_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_gi <- survival::coxph(os_cancer_gi, data = ukb_os)
summary(cancer_gi)

coef_cancer_gi <- summary(cancer_gi)$coef[1, ]
coef_cancer_gi
```

#### Malignant neoplasms of respiratory and intrathoracic organs  
```{r analysis_03 cs_cancer_resp_obj}
cs_cancer_resp_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_resp))
os_cancer_resp <- cs_cancer_resp_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_resp <- survival::coxph(os_cancer_resp, data = ukb_os)
summary(cancer_resp)

coef_cancer_resp <- summary(cancer_resp)$coef[1, ]
coef_cancer_resp
```

#### Malignant neoplasms of bone and articular cartilage  
```{r analysis_03 cs_cancer_bone_obj}
cs_cancer_bone_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_bone))
os_cancer_bone <- cs_cancer_bone_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_bone <- survival::coxph(os_cancer_bone, data = ukb_os)
summary(cancer_bone)

coef_cancer_bone <- summary(cancer_bone)$coef[1, ]
coef_cancer_bone
```

#### Melanoma and other malignant neoplasms of skin  
```{r analysis_03 cs_cancer_skin_obj}
cs_cancer_skin_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_skin))
os_cancer_skin <- cs_cancer_skin_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_skin <- survival::coxph(os_cancer_skin, data = ukb_os)
summary(cancer_skin)

coef_cancer_skin <- summary(cancer_skin)$coef[1, ]
coef_cancer_skin
```

#### Malignant neoplasms of mesothelial and soft tissue  
```{r analysis_03 cs_cancer_soft_tissue_obj}
cs_cancer_soft_tissue_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_soft_tissue))
os_cancer_soft_tissue <- cs_cancer_soft_tissue_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_soft_tissue <- survival::coxph(os_cancer_soft_tissue, data = ukb_os)
summary(cancer_soft_tissue)

coef_cancer_soft_tissue <- summary(cancer_soft_tissue)$coef[1, ]
coef_cancer_soft_tissue
```

#### Malignant neoplasm of breast  
```{r analysis_03 cs_cancer_breast_obj}
cs_cancer_breast_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_breast))
os_cancer_breast <- cs_cancer_breast_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + smk + alc + bmi + wbc + hap 
cancer_breast <- survival::coxph(os_cancer_breast, data = ukb_os, subset = sex == "women")
summary(cancer_breast)

coef_cancer_breast <- summary(cancer_breast)$coef[1, ]
coef_cancer_breast
```

#### Malignant neoplasms of female genital organs  
```{r analysis_03 cs_cancer_female_obj}
cs_cancer_female_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_female))
os_cancer_female <- cs_cancer_female_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + smk + alc + bmi + wbc + hap 
cancer_female <- survival::coxph(os_cancer_female, data = ukb_os, subset = sex == "women")
summary(cancer_female)

coef_cancer_female <- summary(cancer_female)$coef[1, ]
coef_cancer_female
```

#### Malignant neoplasms of male genital organs  
```{r analysis_03 cs_cancer_male_obj}
cs_cancer_male_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_male))
os_cancer_male <- cs_cancer_male_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + smk + alc + bmi + wbc + hap 
cancer_male <- survival::coxph(os_cancer_male, data = ukb_os, subset = sex == "men")
summary(cancer_male)

coef_cancer_male <- summary(cancer_male)$coef[1, ]
coef_cancer_male
```

#### Malignant neoplasms of urinary tract  
```{r analysis_03 cs_cancer_urinary_obj}
cs_cancer_urinary_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_urinary))
os_cancer_urinary <- cs_cancer_urinary_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_urinary <- survival::coxph(os_cancer_urinary, data = ukb_os)
summary(cancer_urinary)

coef_cancer_urinary <- summary(cancer_urinary)$coef[1, ]
coef_cancer_urinary
```

#### Malignant neoplasms of eye, brain, and other parts of central nervous system  
```{r analysis_03 cs_cancer_cns_obj}
cs_cancer_cns_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_cns))
os_cancer_cns <- cs_cancer_cns_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_cns <- survival::coxph(os_cancer_cns, data = ukb_os)
summary(cancer_cns)

coef_cancer_cns <- summary(cancer_cns)$coef[1, ]
coef_cancer_cns
```

#### Malignant neoplasms of thyroid and other endocrine glands  
```{r analysis_03 cs_cancer_endo_obj}
cs_cancer_endo_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_endo))
os_cancer_endo <- cs_cancer_endo_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_endo <- survival::coxph(os_cancer_endo, data = ukb_os)
summary(cancer_endo)

coef_cancer_endo <- summary(cancer_endo)$coef[1, ]
coef_cancer_endo
```

#### Malignant neoplasms of ill-defined, secondary and unspecified sites  
```{r analysis_03 cs_cancer_uns_obj}
cs_cancer_uns_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_uns))
os_cancer_uns <- cs_cancer_uns_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_uns <- survival::coxph(os_cancer_uns, data = ukb_os)
summary(cancer_uns)

coef_cancer_uns <- summary(cancer_uns)$coef[1, ]
coef_cancer_uns
```

#### Hematologic cancer  
```{r analysis_03 cs_cancer_blood_obj}
cs_cancer_blood_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_blood))
os_cancer_blood <- cs_cancer_blood_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_blood <- survival::coxph(os_cancer_blood, data = ukb_os)
summary(cancer_blood)

coef_cancer_blood <- summary(cancer_blood)$coef[1, ]
coef_cancer_blood
```

#### Lymphomas  
```{r analysis_03 cs_cancer_lymphoma_obj}
cs_cancer_lymphoma_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_lymphoma))
os_cancer_lymphoma <- cs_cancer_lymphoma_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_lymphoma <- survival::coxph(os_cancer_lymphoma, data = ukb_os)
summary(cancer_lymphoma)

coef_cancer_lymphoma <- summary(cancer_lymphoma)$coef[1, ]
coef_cancer_lymphoma
```

#### Leukemias  
```{r analysis_03 cs_cancer_leukemia_obj}
cs_cancer_leukemia_obj <- with(ukb_os, Surv(time = time_os, event = cod_cancer_leukemia))
os_cancer_leukemia <- cs_cancer_leukemia_obj ~ mito_score_sum_lc + strata(center) + rcs(age, df= 4) + sex + smk + alc + bmi + wbc + hap 
cancer_leukemia <- survival::coxph(os_cancer_leukemia, data = ukb_os)
summary(cancer_leukemia)

coef_cancer_leukemia <- summary(cancer_leukemia)$coef[1, ]
coef_cancer_leukemia
```

### Combine results   
```{r analysis_03 cod_cancer combine results}
# coefficients 
cod_cancer_coef <- rbind(
  coef_cancer, coef_cancer_blood, coef_cancer_head_neck, coef_cancer_gi, coef_cancer_resp, coef_cancer_bone, coef_cancer_skin, 
  coef_cancer_soft_tissue, coef_cancer_breast, coef_cancer_female, coef_cancer_male, coef_cancer_urinary, coef_cancer_cns, coef_cancer_endo, 
  coef_cancer_uns, coef_cancer_lymphoma, coef_cancer_leukemia
  ) %>%
  as_tibble() %>%
  rename(
    p = `Pr(>|z|)`
  ) 
cod_cancer_coef

# calculate FDR 
cod_cancer_coef %<>%
  mutate(
    fdr = p.adjust(p, "BH"),
    fdr = ifelse(fdr < 0.05, format(fdr, format = "e", digits = 3), round(fdr, 2)),
    # fdr = round(p.adjust(p, "BH"), 2),
    fdr = as.character(fdr)
  ) %>%
  mutate(
    row = c(1:17)
  )
cod_cancer_coef
```

### Merge table with results  
```{r analysis_03 cod_cancer merge table results}
data_cancer <- cod_cancer_table %>%
  left_join(cod_cancer_coef, by = "row") %>%
  mutate(
    # add estimates  
    lci = exp(coef - 1.96*`se(coef)`),
    uci = exp(coef + 1.96*`se(coef)`),
    `HR (95% CI)` = sprintf("%.2f (%.2f to %.2f)", `exp(coef)`, lci, uci),
    ` ` = paste(rep(" ", 25), collapse= " "), 
    # reformat p-values 
    pp = ifelse(p < 0.05, format(p, format = "e", digits = 3), round(p, 2)),
    pp = as.character(pp),
    pp = case_when(
      pp == "1.85e-10" ~ "1.85x10\u207b\u00b9\u2070",
      pp == "4.74e-25" ~ "4.74x10\u207b\u00b2\u2075",
      pp == "1.85e-10" ~ "1.85x10\u207b\u00b9\u2070",
      pp == "1.96e-05" ~ "1.96x10\u207b\u2075",
      pp == "7.72e-04" ~ "7.72x10\u207b\u2074",
      pp == "4.45e-36" ~ "4.45x10\u207b\u00b3\u2076",
      pp == "7.73e-04" ~ "7.73x10\u207b\u2074",
      pp == "2.40e-02" ~ "2.40x10\u207b\u00b2",
      pp == "0.2" ~ "0.20",
      pp == "0.3" ~ "0.30",
      TRUE ~ pp
    )
  ) %>%
  rename(
    `P value` = pp
  ) %>%
  dplyr::select(
    `Cancer type`, Events, ` `, `HR (95% CI)`, `P value`, everything()
  ) %>%
  filter(Events >= 100) %>%
  arrange(-coef) %>%
  mutate(
    `Cancer type` = case_when(
      `Cancer type` == "Lymphoma" ~ "  Lymphoma",
      `Cancer type` == "Leukemia" ~ "  Leukemia",
      TRUE ~ `Cancer type`),
    order = row_number(),
    order = case_when(
      `Cancer type` == "Any cancer" ~ 0,
      `Cancer type` == "Hematologic" ~ 1,
      `Cancer type` == "  Leukemia" ~ 2,
      `Cancer type` == "  Lymphoma" ~ 3,
      TRUE ~  as.numeric(order)
    )
  ) %>%
  arrange(order)
data_cancer
```

### Figure 6b  
```{r analysis_03 figure6b, fig.height=4.5, fig.width=9}
fig6b <- forestploter::forest(
  data_cancer[, c(1:5)],
  est = data_cancer$`exp(coef)`,
  lower = data_cancer$lci,
  upper = data_cancer$uci,
  sizes = 0.4,
  ci_column = 3,
  ref_line = 1,
  xlog = TRUE,
  xlim = c(0.5, 6), 
  ticks_at = c(0.5, 0.75, 1, 1.5, 2.0, 3, 4, 6),
  title = "Hazard ratios for each cause of death"
  )
fig6b
```

# Figure 6  
```{r analysis_03 figure6, fig.height=11, fig.width=9}
fig6 <- ggpubr::ggarrange(
  fig6a, fig6b,
  ncol = 1, nrow = 2,
  labels = c("a", "b"),
  align = "v"
)
fig6

ggsave("ukb_heteroplasmy_outcomes_fig6.tiff", plot = fig6, height=11, width=9) 
```


