---
title: "UKB - Heteroplasmy and mortality - data cleaning"
subtitle: "002 - mortality outcomes"
author: "Yun Soo Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    pandoc_args: [
    "--number-sections",
    "--number-offset=0"
    ]
fontsize: 11pt
---

# Set up  

## Program description  
* This program reads in two mortality outcome data sets downloaded from the UK Biobank RAP on June 8, 2022 - 1) death records (`mt_heteroplasmy_death_record_death_2022_06_08.csv`) and 2) cause of death (`mt_heteroplasmy_death_cause_record_death_cause_2022_06_08.csv`) -  cleans the raw data, and saves the object into an .rds ("ukb_heteroplasmy_mortality_02_mortality.rds").   
  * Step 1: Read in data  
  * Step 2: Multiple records of death    
  * Step 3: Merge data sets   
  * Step 4: Generate new variables for each cause of death  
  * Step 5: Save data   
* Notes: In the case when there is more than 1 death records per person, we use observations identified as `ins_index` == 0. In addition, we use only the primary cause of death (`level` == 1) for this study. In the exploratory phase, we have determined that the date of death is identical in the two death records.   

## Versions of R and packages  
```{r clean_02 set up, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r clean_02 load libraries, echo = TRUE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(stringr)
library(data.table)
```

## Print out settings  
```{r clean_02 print settings}
options(tibble.print_max = 100)
options(pillar.sigfig = 5)
```

# Load data  
## Death  
* 60 participants have 2 primary causes of death (`ins_index` == 0, `ins_index` == 1). This may happen when there is more than one sources of death registry     

```{r clean_02 load death data}
death <- fread("../../data/mt_heteroplasmy_death_record_death_2022_06_08.csv") %>%
  as.data.frame()
death

# 37,897 participants, 37,957 observations  
unique(death$eid) %>% 
  length()

# index of death information
table(death$ins_index)
```

## Cause of death  
* Each individual may have more than one source of death registry (`ins_index`) -- up to 2 sources    
* Each individual may have more than one cause of death (`arr_index`) -- up to 15 causes of death    
  * `level` indicates whether the cause of death is primary (`level` == 1) or secondary (`level` == 2)  
* In this case, we use `ins_index` == 0 and use its primary cause of death as the outcome  

```{r clean_02 load cod data}
cod <- fread("../../data/mt_heteroplasmy_death_cause_record_death_cause_2022_06_08.csv") %>%
  as.data.frame()
cod

# 37,731 participants, 94,440 observations  
unique(cod$eid) %>%
  length()

# death registry data  
table(cod$ins_index)

# cause of death  
table(cod$arr_index)

# primary or secondary causes of death 
table(cod$level)
```

# Multiple records of death  
## Death  
* Confirm if the date of death is the same from the 2 distinct records -> all date of death records are identical for each individual    

```{r clean_02 death multiple}
death %>%
  # select participants with 2 or more death records
  group_by(eid) %>%
  mutate(nrows = n()) %>%
  ungroup() %>%
  filter(nrows != 1) %>%  
  # select relevant variables  
  dplyr::select(eid, ins_index, date_of_death) %>%
  # long to wide format 
  pivot_wider(names_from = ins_index,
              names_glue = "ins_{ins_index}",
              values_from = date_of_death) %>%
  # difference between the first and second record
  mutate(diff = difftime(ins_0, ins_1, units = "days"),
         diff = as.numeric(diff)) %>%
  filter(diff != 0) # 0 observation indicating that date of death from the 2 data sets match  
```

## Cause of death  
* The primary cause of death may differ between the two sources of death information  
* In many cases, the primary cause of death from one source is included as a secondary cause of death from the other  

```{r clean_02 cod multiple}
cod %>%
  group_by(eid) %>%
  mutate(nins = max(ins_index)) %>%
  ungroup() %>%
  filter(nins != 0) %>%
  dplyr::select(eid, ins_index, arr_index, level, cause_icd10) %>%
  arrange(eid, ins_index, arr_index)

cod %>% 
  janitor::tabyl(level, arr_index)
```

# Merge data sets 
## Keep `ins_index` == 0 and its primary cause of death  
### Death  
* 37897 observations with the same number of participants  

```{r clean_02 death select ins_index}
death %<>%
  filter(ins_index == 0) %>%
  dplyr::select(eid, dsource, date_of_death)

death %>% 
  summarise(
    n = n_distinct(eid),
    nrows = n()
  )
```

### Cause of death  
* 37731 observations with the same number of participants  
* There are 166 (37897 - 37731) participants without an identified primary cause of death but we still include them in the final data set.   

```{r clean_02 cod select ins_index}
cod %<>%
  # use only one of the 2 death records and primary cause of death
  filter(ins_index == 0 & level == 1) %>%
  dplyr::select(eid, cause_icd10)

cod %>% 
  summarise(
    n = n_distinct(eid),
    nrows = n()
  )
```

## Merge two data sets  
* Merge data sets by `eid` and `ins_index`  
* As `death` data frame has all the individuals who died (including the ones without information on the cause of death, n = 37897), merge `cod` data frame to `death` data frame. Otherwise, we end up having different numbers of death cases in the analysis.  
  * 37897 observations with the same number of participants  
* The last date of death in the data is "2021-11-12" and we use this date as the date of administrative censoring.    

```{r clean_02 merge}
death_cause <- death %>%
  left_join(cod, by = "eid") %>%
  arrange(eid)
glimpse(death_cause)

death_cause %>% 
  summarise(
    n = n_distinct(eid),
    nrows = n()
  )

death_cause %>%
  count(cause_icd10 == "") # n = 166 without cause of death 

# last date of follow-up in the death data = date of administrative censoring  
max(death_cause$date_of_death) # the last date of death identified in the data is "2021-11-12"
```

# Generate new variables for each cause of death  
* Infections: A00-B99, L00-L08
* Neoplasms: C00-D48  
  * Cancers: C00-C97  
    * Solid cancers: C00-C80  
    * Hematologic cancers: C81-C96
  * Benign neoplasms: D00-D48  
    * Polycythemia vera: D45  
    * Myelodysplastic syndrome: D46  
* Benign diseases of the blood: D50-D89  
* Endocrine disorders: E00-E90  
* Mental and behavioral disorders: F00-F89
* Neurological disorders: G00-G99  
  * Neurodegenerative diseases: G30, G31  
* Eye and adnexa disorders: H00-H59 - there is 0 participants  
* Ear and mastoid process disorders: H60-H95 (H65-H75) 
* Circulatory disorders: I00-I99  
  * Acute myocardial infarction: I21  
  * Chronic ischemic heart disease: I25  
  * Cerebrovascular disease: I60-I69  
* Respiratory disorders: J00-J99  
  * Chronic lower respiratory disease: J40-J47  
* Digestive disorders: K00-K95  
  * Upper GI: K20-K31  
  * Lower GI: K50-K64  
  * Liver diseases: K70-K77  
* Diseases of the skin and subcutaneous tissue: L10-L99
* Diseases of the musculoskeletal system and connective tissue: M00-M99
* Genitourinary disorders: N00-N99    
* Pregnancy, childbirth and the puerperium: O00-O9A
* Congenital malformations, deformations and chromosomal anomalies: Q00-Q99  
* Symptoms, signs, and abnormal clinical and laboratory findings, not elsewhere classified: R00-R99  
* Codes for special purposes: U00-U85
  * COVID-19: U07  
* External causes: V01-Y89  
  * Accidents: V01-V59  
  * Intentional self-harm: X60-X84  

## Label settings  
```{r clean_02 label}
noyes <- c("no", "yes")
```

## All-cause mortality  
* All participants included in this data set is coded as `death = 1`
```{r clean_02 all-cause}
death_cause %<>%
  mutate(death = 1)
```

## Infections: A00-B99, L00-L08  

```{r clean_02 infection}
death_cause %<>%
  mutate(
    cod_infection = 
      case_when(
        grepl("A", cause_icd10) ~ 1,
        grepl("B", cause_icd10) ~ 1,
        grepl("L0", cause_icd10) ~ 1,
        TRUE ~ 0),
    cod_infection =
      factor(
        cod_infection, 
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_infection, useNA = "ifany")
```

## Neoplasms: C00-D48  
```{r clean_02 neoplasm}
benign_neoplasm <- c("D1", "D2", "D3", "D4")

death_cause %<>%
  mutate(
    cod_neoplasm = 
      case_when(
        grepl("C", cause_icd10) ~ 1,
        grepl(paste(benign_neoplasm, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_neoplasm =
      factor(
        cod_neoplasm,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_neoplasm, useNA = "ifany")
```

### Cancer: C00-C97    
```{r clean_02 cancer}
death_cause %<>%
  mutate(
    cod_cancer = 
      case_when(
        grepl("C", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer = 
      factor(
        cod_cancer,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer, useNA = "ifany")
```

#### Solid cancers: C00-C80    
```{r clean_02 cancer_solid}
cancer_solid <- c("C0", "C1", "C2", "C3", "C4", "C5", "C6", "C7", "C80")

death_cause %<>%
  mutate(
    cod_cancer_solid = 
      case_when(
        grepl(paste(cancer_solid, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_solid = 
      factor(
        cod_cancer_solid,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_solid, useNA = "ifany")
```

#### Hematologic cancers: C81-C96  
```{r clean_02 cancer_blood}
cancer_blood <- c("C81", "C82", "C83", "C84", "C85", "C86", "C87", "C88", "C89", 
                  "C90", "C91", "C92", "C93", "C94", "C95", "C96")

death_cause %<>%
  mutate(
    cod_cancer_blood = 
      case_when(
        grepl(paste(cancer_blood, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_blood = 
      factor(
        cod_cancer_blood,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_blood, useNA = "ifany")
```

#### Malignant neoplasms of lip, oral cavity, and pharynx
* ICD10 C00-C14   
```{r clean_02 cancer_head_neck}
cancer_head_neck <- c("C0", "C10", "C11", "C12", "C13", "C14")

death_cause %<>%
  mutate(
    cod_cancer_head_neck = 
      case_when(
        grepl(paste(cancer_head_neck, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_head_neck = 
      factor(
        cod_cancer_head_neck,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_head_neck, useNA = "ifany")
```

#### Malignant neoplasms of digestive organs
* ICD10 C15-C26  
```{r clean_02 cancer_gi}
cancer_gi <- c("C15", "C16", "C17", "C18", "C19", "C20", "C21", "C22", "C23", "C24", "C25", "C26")

death_cause %<>%
  mutate(
    cod_cancer_gi = 
      case_when(
        grepl(paste(cancer_gi, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_gi = 
      factor(
        cod_cancer_gi,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_gi, useNA = "ifany")
```

#### Malignant neoplasms of respiratory and intrathoracic organs  
* ICD10 C30-C39  
```{r clean_02 cancer_resp}
cancer_resp <- c("C30", "C31", "C32", "C33", "C34", "C37", "C38", "C39")

death_cause %<>%
  mutate(
    cod_cancer_resp = 
      case_when(
        grepl(paste(cancer_resp, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_resp = 
      factor(
        cod_cancer_resp,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_resp, useNA = "ifany")
```

#### Malignant neoplasms of bone and articular cartilage    
* ICD10 C40-C41  
```{r clean_02 cancer_bone}
cancer_bone <- c("C40", "C41")

death_cause %<>%
  mutate(
    cod_cancer_bone = 
      case_when(
        grepl(paste(cancer_bone, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_bone = 
      factor(
        cod_cancer_bone,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_bone, useNA = "ifany")
```

#### Melanoma and other malignant neoplasms of skin  
* ICD10 C43
* Non-melanoma skin cancer (C44) is not considered as a cancer outcome    
```{r clean_02 cancer_skin}
death_cause %<>%
  mutate(
    cod_cancer_skin = 
      case_when(
        grepl("C43", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_skin = 
      factor(
        cod_cancer_skin,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_skin, useNA = "ifany")
```

#### Malignant neoplasms of mesothelial and soft tissue  
* ICD10 C45-C49  
```{r clean_02 cancer_soft_tissue}
cancer_soft_tissue <- c("C45", "C46", "C47", "C48", "C49")

death_cause %<>%
  mutate(
    cod_cancer_soft_tissue = 
      case_when(
        grepl(paste(cancer_soft_tissue, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_soft_tissue = 
      factor(
        cod_cancer_soft_tissue,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_soft_tissue, useNA = "ifany")
```

#### Malignant neoplasm of breast  
* ICD10 C50  
```{r clean_02 cancer_breast}
death_cause %<>%
  mutate(
    cod_cancer_breast = 
      case_when(
        grepl("C50", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_breast = 
      factor(
        cod_cancer_breast,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_breast, useNA = "ifany")
```

#### Malignant neoplasms of female genital organs  
* ICD10 C51-C58  
```{r clean_02 cancer_female}
cancer_female <- c("C51", "C52", "C53", "C54", "C55", "C56", "C57", "C58")

death_cause %<>%
  mutate(
    cod_cancer_female = 
      case_when(
        grepl(paste(cancer_female, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_female = 
      factor(
        cod_cancer_female,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_female, useNA = "ifany")
```

#### Malignant neoplasms of male genital organs  
* ICD10 C60-C63
```{r clean_02 cancer_male}
cancer_male <- c("C60", "C61", "C62", "C63")

death_cause %<>%
  mutate(
    cod_cancer_male = 
      case_when(
        grepl(paste(cancer_male, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_male = 
      factor(
        cod_cancer_male,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_male, useNA = "ifany")
```

#### Malignant neoplasms of urinary tract  
* ICD10 C64-C68  
```{r clean_02 cancer_urinary}
cancer_urinary <- c("C64", "C65", "C66", "C67", "C68")

death_cause %<>%
  mutate(
    cod_cancer_urinary = 
      case_when(
        grepl(paste(cancer_urinary, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_urinary = 
      factor(
        cod_cancer_urinary,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_urinary, useNA = "ifany")
```

#### Malignant neoplasms of eye, brain, and other parts of central nervous system  
* ICD10 C69-C72  
```{r clean_02 cancer_cns}
cancer_cns <- c("C69", "C70", "C71", "C72")

death_cause %<>%
  mutate(
    cod_cancer_cns = 
      case_when(
        grepl(paste(cancer_cns, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_cns = 
      factor(
        cod_cancer_cns,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_cns, useNA = "ifany")
```

#### Malignant neoplasms of thyroid and other endocrine glands  
* ICD10 C73-C75  
```{r clean_02 cancer_endo}
cancer_endo <- c("C73", "C74", "C75")

death_cause %<>%
  mutate(
    cod_cancer_endo = 
      case_when(
        grepl(paste(cancer_endo, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_endo = 
      factor(
        cod_cancer_endo,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_endo, useNA = "ifany")
```

#### Malignant neoplasms of ill-defined, secondary and unspecified sites  
* ICD10 C76-C80  
```{r clean_02 cancer_uns}
cancer_uns <- c("C76", "C77", "C78", "C79", "C80")

death_cause %<>%
  mutate(
    cod_cancer_uns = 
      case_when(
        grepl(paste(cancer_uns, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_uns = 
      factor(
        cod_cancer_uns,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_uns, useNA = "ifany")
```

#### Lymphomas  
* ICD10 C81-C86
```{r clean_02 cancer_lymphoma}
cancer_lymphoma <- c("C81", "C82", "C83", "C84", "C85", "C86")

death_cause %<>%
  mutate(
    cod_cancer_lymphoma = 
      case_when(
        grepl(paste(cancer_lymphoma, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_lymphoma = 
      factor(
        cod_cancer_lymphoma,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_lymphoma, useNA = "ifany")
```

#### Malignant immunoproliferative diseases  
* ICD10 C88
```{r clean_02 cancer_ipd}
death_cause %<>%
  mutate(
    cod_cancer_ipd = 
      case_when(
        grepl("C88", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_ipd = 
      factor(
        cod_cancer_ipd,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_ipd, useNA = "ifany")
```

#### Multiple myeloma and malignant plasma cell neoplasms  
* ICD10 C90
```{r clean_02 cancer_mm}
death_cause %<>%
  mutate(
    cod_cancer_mm = 
      case_when(
        grepl("C90", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_mm = 
      factor(
        cod_cancer_mm,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_mm, useNA = "ifany")
```

#### Leukemias  
* ICD10 C91-C95
```{r clean_02 cancer_leukemia}
cancer_leukemia <- c("C91", "C92", "C93", "C94", "C95")

death_cause %<>%
  mutate(
    cod_cancer_leukemia = 
      case_when(
        grepl(paste(cancer_leukemia, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_leukemia = 
      factor(
        cod_cancer_leukemia,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_leukemia, useNA = "ifany")
```

#### Other and unspecified malignant neoplasms of lymphoid, hematopoietic and related tissue  
* ICD10 C96  
```{r clean_02 cancer_blood_uns}
death_cause %<>%
  mutate(
    cod_cancer_blood_uns = 
      case_when(
        grepl("C96", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_blood_uns = 
      factor(
        cod_cancer_blood_uns,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_blood_uns, useNA = "ifany")
```

#### Malignant neoplasms of independent (primary) multiple sites  
* ICD10 C97
```{r clean_02 cancer_multiple}
death_cause %<>%
  mutate(
    cod_cancer_multiple = 
      case_when(
        grepl("C97", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cancer_multiple = 
      factor(
        cod_cancer_multiple,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_cancer_multiple, useNA = "ifany")
```

### Benign neoplasms 
```{r clean_02 benign neoplasms}
death_cause %<>%
  mutate(
    cod_benign_neoplasm = 
      case_when(
        grepl(paste(benign_neoplasm, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_benign_neoplasm = 
      factor(
        cod_benign_neoplasm,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_benign_neoplasm, useNA = "ifany")
```

#### Polycythemia vera  
* ICD10 D45  
```{r clean_02 pv}
death_cause %<>%
  mutate(
    cod_pv = 
      case_when(
        grepl("D45", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_pv = 
      factor(
        cod_pv,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_pv, useNA = "ifany")
```

#### Myelodysplastic syndromes  
* ICD10 D46  
```{r clean_02 mds}
death_cause %<>%
  mutate(
    cod_mds = 
      case_when(
        grepl("D46", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_mds = 
      factor(
        cod_mds,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_mds, useNA = "ifany")
```

#### Other neoplasms of uncertain or unknown behavior of lymphoid, hematopoietic and related tissue
* ICD10 D47  
```{r clean_02 unknown_blood}
death_cause %<>%
  mutate(
    cod_unknown_blood = 
      case_when(
        grepl("D47", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_unknown_blood = 
      factor(
        cod_unknown_blood,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_unknown_blood, useNA = "ifany")
```

## Benign diseases of the blood: D50-D89  
```{r clean_02 benign blood}
benign_blood <- c("D5", "D6", "D7", "D8")

death_cause %<>%
  mutate(
    cod_blood = 
      case_when(
        grepl(paste(benign_blood, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_blood = 
      factor(
        cod_blood,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_blood, useNA = "ifany")
```

## Endocrine disorders: E00-E90  
```{r clean_02 endocrine}
death_cause %<>%
  mutate(
    cod_endocrine = 
      case_when(
        grepl("E", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_endocrine =
      factor(
        cod_endocrine, 
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_endocrine, useNA = "ifany")
```

## Mental and behavioral disorders: F00-F89  
```{r clean_02 mental}
death_cause %<>%
  mutate(
    cod_mental = 
      case_when(
        grepl("F", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_mental = 
      factor(
        cod_mental,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_mental, useNA = "ifany")
```

## Neurological disorders: G00-G99  
```{r clean_02 neuro}
death_cause %<>%
  mutate(
    cod_neuro = 
      case_when(
        grepl("G", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_neuro = 
      factor(
        cod_neuro,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_neuro, useNA = "ifany")
```

### Neurodenegerative diseases: G30, G31  
```{r clean_02 neuro_nds}
death_cause %<>%
  mutate(
    cod_neuro_nds = 
      case_when(
        grepl("G30", cause_icd10) ~ 1,
        grepl("G31", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_neuro_nds = 
      factor(
        cod_neuro_nds,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_neuro_nds, useNA = "ifany")
```

## Eye and adnexa disorders: H00-H59
```{r clean_02 eye}
eye <- c("H0", "H1", "H2", "H3", "H4", "H5")

death_cause %<>%
  mutate(
    cod_eye = 
      case_when(
        grepl(paste(eye, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_eye = 
      factor(
        cod_eye,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_eye, useNA = "ifany")
```

## Ear and mastoid process disorders: H60-H95
```{r clean_02 ear}
ear <- c("H6", "H7", "H8", "H9")

death_cause %<>%
  mutate(
    cod_ear = 
      case_when(
        grepl(paste(ear, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_ear = 
      factor(
        cod_ear,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_ear, useNA = "ifany")
```

## Circulatory disorders: I00-I99  
```{r clean_02 cardio}
death_cause %<>%
  mutate(
    cod_cardio = 
      case_when(
        grepl("I", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cardio = 
      factor(
        cod_cardio,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_cardio, useNA = "ifany")
```

### Acute myocardial infarction  
```{r clean_02 cardio_ami}
death_cause %<>%
  mutate(
    cod_cardio_ami = 
      case_when(
        grepl("I21", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cardio_ami = 
      factor(
        cod_cardio_ami,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_cardio_ami, useNA = "ifany")
```

### Chronic ischemic heart disease  
```{r clean_02 cardio_cihd}
death_cause %<>%
  mutate(
    cod_cardio_cihd = 
      case_when(
        grepl("I25", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cardio_cihd = 
      factor(
        cod_cardio_cihd,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_cardio_cihd, useNA = "ifany")
```

### Cerebrovascular disease  
```{r clean_02 cardio_cv}
death_cause %<>%
  mutate(
    cod_cardio_cv = 
      case_when(
        grepl("I6", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_cardio_cv = 
      factor(
        cod_cardio_cv,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_cardio_cv, useNA = "ifany")
```

## Respiratory disorders: J00-J99  
```{r clean_02 pulmo}
death_cause %<>%
  mutate(
    cod_pulmo = 
      case_when(
        grepl("J", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_pulmo = 
      factor(
        cod_pulmo,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_pulmo, useNA = "ifany")
```

### Chronic lower respiratory disease  
```{r clean_02 pulmo_clrd}
death_cause %<>%
  mutate(
    cod_pulmo_clrd = 
      case_when(
        grepl("J4", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_pulmo_clrd = 
      factor(
        cod_pulmo_clrd,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_pulmo_clrd, useNA = "ifany")
```

## Digestive disorders: K00-K95  
```{r clean_02 gi}
death_cause %<>%
  mutate(
    cod_gi = 
      case_when(
        grepl("K", cause_icd10) ~ 1,
        TRUE ~ 0
      ), 
    cod_gi = 
      factor(
        cod_gi, 
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_gi, useNA = "ifany")
```

### Upper GI diseases  
```{r clean_02 gi_ugi}
death_cause %<>%
  mutate(
    cod_gi_ugi = 
      case_when(
        grepl("K2", cause_icd10) ~ 1,
        grepl("K31", cause_icd10) ~ 1, 
        TRUE ~ 0
      ), 
    cod_gi_ugi = 
      factor(
        cod_gi_ugi, 
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_gi_ugi, useNA = "ifany")
```

### Lower GI
```{r clean_02 gi_lgi}
death_cause %<>%
  mutate(
    cod_gi_lgi = 
      case_when(
        grepl("K5", cause_icd10) ~ 1,
        grepl("K61", cause_icd10) ~ 1, 
        grepl("K62", cause_icd10) ~ 1, 
        grepl("K63", cause_icd10) ~ 1, 
        TRUE ~ 0
      ), 
    cod_gi_lgi = 
      factor(
        cod_gi_lgi, 
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_gi_lgi, useNA = "ifany")
```

### Liver diseases  
```{r clean_02 gi_liver}
death_cause %<>%
  mutate(
    cod_gi_liver = 
      case_when(
        grepl("K7", cause_icd10) ~ 1,
        TRUE ~ 0
      ), 
    cod_gi_liver = 
      factor(
        cod_gi_liver, 
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_gi_liver, useNA = "ifany")
```

## Diseases of the skin and subcutaneous tissue: L10-L99  
```{r clean_02 skin}
skin <- c("L1", "L2", "L3", "L4", "L5", "L6", "L7", "L8", "L9")

death_cause %<>%
  mutate(
    cod_skin = 
      case_when(
        grepl(paste(skin, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_skin = 
      factor(
        cod_skin,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_skin, useNA = "ifany")
```

## Diseases of the musculoskeletal system and connective tissue: M00-M99  
```{r clean_02 muscle}
death_cause %<>%
  mutate(
    cod_muscle = 
      case_when(
        grepl("M", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_muscle = 
      factor(
        cod_muscle,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_muscle, useNA = "ifany")
```

## Genitourinary disorders: N00-N99   
```{r clean_02 gu}
death_cause %<>%
  mutate(
    cod_gu = 
      case_when(
        grepl("N", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_gu = 
      factor(
        cod_gu,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_gu, useNA = "ifany")
```

## Pregnancy, childbirth and the puerperium: O00-O9A  
```{r clean_02 pregnancy}
death_cause %<>%
  mutate(
    cod_pregnancy = 
      case_when(
        grepl("O", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_pregnancy = 
      factor(
        cod_pregnancy,
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_pregnancy, useNA = "ifany")
```

## Congenital malformations, deformations and chromosomal anomalies: Q00-Q99   
```{r clean_02 congenital}
death_cause %<>%
  mutate(
    cod_congenital = 
      case_when(
        grepl("Q", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_congenital = 
      factor(
        cod_congenital, 
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_congenital, useNA = "ifany")
```

## Symptoms, signs, and abnormal clinical and laboratory findings, not elsewhere classified: R00-R99  
```{r clean_02 uns}
death_cause %<>%
  mutate(
    cod_uns = 
      case_when(
        grepl("R", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_uns = 
      factor(
        cod_uns, 
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_uns, useNA = "ifany")
```

## Codes for special purposes: U00-U85  
```{r clean_02 special}
death_cause %<>%
  mutate(
    cod_special = 
      case_when(
        grepl("U", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_special = 
      factor(
        cod_special, 
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_special, useNA = "ifany")
```

### COVID-19  
```{r clean_02 covid}
death_cause %<>%
  mutate(
    cod_covid = 
      case_when(
        grepl("U07", cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_covid = 
      factor(
        cod_covid, 
        levels = c(0, 1),
        labels = noyes
      )
  )

table(death_cause$cod_covid, useNA = "ifany")
```

## External causes: V01-Y89    
```{r clean_02 external}
death_cause %<>%
  mutate(
    cod_external = 
      case_when(
        grepl("V", cause_icd10) ~ 1,
        grepl("W", cause_icd10) ~ 1,
        grepl("X", cause_icd10) ~ 1,
        grepl("Y", cause_icd10) ~ 1, 
        TRUE ~ 0
      ),
    cod_external = 
      factor(
        cod_external,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_external, useNA = "ifany")
```

### Accidents  
```{r clean_02 external_accident}
external_accident <- c("V0", "V1", "V2", "V3", "V4", "V5")

death_cause %<>%
  mutate(
    cod_external_accident = 
      case_when(
        grepl(paste(external_accident, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_external_accident = 
      factor(
        cod_external_accident,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_external_accident, useNA = "ifany")
```

### Intentional self-harm  
```{r clean_02 external_selfharm}
external_selfharm <- c("X6", "X7", "X80", "X81", "X82", "X83", "X84")

death_cause %<>%
  mutate(
    cod_external_selfharm = 
      case_when(
        grepl(paste(external_selfharm, collapse = "|"), cause_icd10) ~ 1,
        TRUE ~ 0
      ),
    cod_external_selfharm = 
      factor(
        cod_external_selfharm,
        levels = c(0, 1), 
        labels = noyes
      )
  )

table(death_cause$cod_external_selfharm, useNA = "ifany")
```

# Save data  
* Long to wide format with cause of death variables as each variable  
```{r clean_02 save}
death_cause %<>%
  mutate(eid = as.character(eid)) %>%
  rename(id = eid)

death_cause %>%
  skimr::skim_without_charts()

death_cause %>%
  saveRDS("ukb_heteroplasmy_mortality_02_mortality.rds")
```


