---
title: "UKB - Heteroplasmy and mortality - data cleaning"
subtitle: "001 - baseline characteristics"
author: "Yun Soo Hong"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document:
    toc: true
    toc_float: 
      collapsed: false
      smooth_scroll: false
    pandoc_args: [
    "--number-sections",
    "--number-offset=0"
    ]
fontsize: 11pt
---

# Set up  
* This program reads in the data set of baseline characteristics downloaded from the UK Biobank RAP on June 18, 2022  (`mt_heteroplasmy_baseline_participant_2022_06_18.csv`), cleans the raw data, and saves the object into an .rds ("ukb_heteroplasmy_mortality_01_baseline.rds"). For categorical variables, those ending with `_na` use `NA` for missing values.    
  * Step 1: Read in data  
  * Step 2: Rename variables
    * After renaming, remove unused variables  
  * Step 3: Generate new variables    
  * Step 4: Save data set  
  
## Versions of R and packages  
```{r clean_01 set up, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries
```{r clean_01 load libraries, echo = TRUE, message = FALSE}
library(tidyverse)
library(magrittr)
library(lubridate)
library(stringr)
library(data.table)
```

## Print out settings  
```{r clean_01 print settings}
options(tibble.print_max = 100)
options(pillar.sigfig = 5)
```

# Load data  
```{r clean_01 load data}
baseline <- fread("../../data/mt_heteroplasmy_baseline_participant_2022_06_18.csv") %>%
  as.data.frame()
baseline
```

# Rename variables  
## Participant ID
```{r clean_01 id}
baseline %<>%
  mutate(eid = as.character(eid)) %>%
  rename(id = eid)
```

## Age at baseline  
```{r clean_01 age}
baseline %<>%
  rename(age = `21022-0.0`)
```

## Year and month of birth  
```{r clean_01 birth}
baseline %<>%
  rename(yob = `34-0.0`,
         mob = `52-0.0`)
```

## Sex  
```{r clean_01 sex}
table(baseline$`31-0.0`, useNA = "ifany")

baseline %<>%
  rename(sex = `31-0.0`) %>%
  mutate(sex = 
           case_when(
             sex == 0 ~ "women", 
             sex == 1 ~ "men"),
         sex = 
           factor(
             sex, 
             levels = c("women", "men"),
             labels = c("women", "men"))
         )
```

## Race / ethnicity  
```{r clean_01 ethnicity}
baseline %<>%
  rename(ethnicity = `21000-0.0`)
table(baseline$ethnicity, useNA = "ifany")

baseline %<>%
  mutate(
    ethnicity_new = case_when(
      ethnicity == 1 | ethnicity == 1001 | ethnicity == 1002 | ethnicity == 1003 ~ 1,
      ethnicity == 2 | ethnicity == 2001 | ethnicity == 2002 | ethnicity == 2003 | ethnicity == 2004 ~ 2,
      ethnicity == 3 | ethnicity == 3001 | ethnicity == 3002 | ethnicity == 3003 | ethnicity == 3004 ~ 3,
      ethnicity == 4 | ethnicity == 4001 | ethnicity == 4002 | ethnicity == 4003 ~ 4,
      ethnicity == 5 ~ 5,
      ethnicity == 6 ~ 6,
      TRUE ~ as.numeric(NA)
    ),
    race = factor(
      ethnicity_new,
      levels = c(1, 2, 3, 4, 5, 6),
      labels = c("white", "mixed", "asian", "black", "chinese", "others")
    ),
    race_na = fct_recode(
      race,
      "white" = "white", 
      "black" = "black", 
      "others" = "others", 
      "others" = "mixed", 
      "asian" = "asian", 
      "asian" = "chinese"
    ),
    race_na = factor(
      race_na, 
      levels = c("white", "black", "asian", "others")
    ),
    race_new = 
      fct_explicit_na(
        race_na, na_level = "missing"
      )
  ) 

table(baseline$race, useNA = "ifany")
table(baseline$race_na, useNA = "ifany")
table(baseline$race_new, useNA = "ifany")
```

## Date of visit  
```{r clean_01 dov}
baseline %<>%
  rename(dov = `53-0.0`) %>%
  mutate(dov = as.Date(dov))
```

## Date and time of blood collection  
```{r clean_01 dobl}
baseline %<>%
  rename(dobl = `3166-0.0`) %>%
  mutate(dobl = as.Date(dobl))
```

## Center  
```{r clean_01 center}
baseline %<>%
  rename(center = `54-0.0`) %>%
  mutate(center = factor(center))
```

## Current employment  
* Use 2 variables: `6142-0.0` and `20119-0.0`  

```{r clean_01 current employment}
table(baseline$`6142-0.0`, useNA = "ifany")

baseline %<>%
  rename(cur_emp = `6142-0.0`) %>%
  mutate(cur_emp = 
           case_when(
             cur_emp == "-7" ~ "none of the above",
             cur_emp == "-3" ~ "prefer not to answer",
             cur_emp == "1" ~ "paid employment",
             cur_emp == "2" ~ "retired",
             cur_emp == "3" ~ "looking after home/family",
             cur_emp == "4" ~ "unable to work", 
             cur_emp == "5" ~ "unemployed", 
             cur_emp == "6" ~ "unpaid or voluntary work", 
             cur_emp == "7" ~ "student",
             # if participant answered more than 1, we select the one coded using the smallest value
             grepl("1", cur_emp) ~ "paid employment",
             grepl("2", cur_emp) ~ "retired",
             grepl("3", cur_emp) ~ "looking after home/family",
             grepl("4", cur_emp) ~ "unable to work",
             grepl("5", cur_emp) ~ "unemployed",
             grepl("6", cur_emp) ~ "unpaid or voluntary work", 
             grepl("7", cur_emp) ~ "student")
         )

baseline %>%
  janitor::tabyl(cur_emp)

# update the current employment status using the `20119-0.0` ("Current employment status (corrected)")
table(baseline$`20119-0.0`, useNA = "ifany")

baseline %<>%
  mutate(cur_emp =
           case_when(
             cur_emp == "none of the above"  | `20119-0.0` == "-7" ~ "none of the above",
             cur_emp == "prefer not to answer" | `20119-0.0` == "-3" ~ "prefer not to answer",
             cur_emp == "paid employment" ~ "paid employment", 
             cur_emp == "retired" | `20119-0.0` == "2" ~ "retired",
             cur_emp == "looking after home/family" | `20119-0.0` == "3" ~ "looking after home/family",
             cur_emp == "unable to work" | `20119-0.0` == "4" ~ "unable to work",
             cur_emp == "unemployed" | `20119-0.0` == "5" ~ "unemployed",
             cur_emp == "unpaid or voluntary work" | `20119-0.0` == "6" ~ "unpaid or voluntary work",
             cur_emp == "student" | `20119-0.0` == "7" ~ "student"
           ),
         cur_emp = 
           factor(
             cur_emp,
             levels = c("paid employment", "retired", "looking after home/family", "unable to work",
                        "unemployed", "unpaid or voluntary work", "student", "none of the above", "prefer not to answer"),
             labels = c("paid employment", "retired", "looking after home/family", "unable to work",
                        "unemployed", "unpaid or voluntary work", "student", "none of the above", "prefer not to answer"))
         ) %>%
  dplyr::select(-`20119-0.0`)

baseline %>%
  janitor::tabyl(cur_emp)

# collapsed version of cur_emp
# people who did not answer or responded "prefer not to answer" will be coded as missing 
baseline %<>%
  mutate(cur_emp_na = 
           fct_recode(
             cur_emp,
             "yes" = "paid employment",
             "yes" = "unpaid or voluntary work", 
             "no" = "retired",
             "no" = "looking after home/family",
             "no" = "unable to work",
             "no" = "unemployed",
             "no" = "student", 
             "none of the above" = "none of the above"
           ),
         cur_emp_na = 
           factor(
             cur_emp_na,
             levels = c("no", "yes", "none of the above"),
             labels = c("no", "yes", "none of the above")
           ),
         cur_emp_col = 
           fct_explicit_na(
             cur_emp_na,
             na_level = "missing")
         ) %>%
  dplyr::select(id, age, yob, mob, sex, cur_emp, cur_emp_na, cur_emp_col, everything())

table(baseline$cur_emp_na, useNA = "ifany")
table(baseline$cur_emp_col, useNA = "ifany")
```

## Education level   
```{r clean_01 education}
table(baseline$`6138-0.0`, useNA = "ifany")

baseline %<>%
  rename(edu_level = `6138-0.0`) %>%
  mutate(edu_level = 
           case_when(
             edu_level == "-7" ~ "none of the above",
             edu_level == "-3" ~ "prefer not to answer",
             edu_level == "1" ~ "college or university degree",
             edu_level == "2" ~ "A or AS levels",
             edu_level == "3" ~ "O level or GCSEs",
             edu_level == "4" ~ "CSEs", 
             edu_level == "5" ~ "NVQ or HND or HNC", 
             edu_level == "6" ~ "other professional qualifications",
             # if participant answered more than 1, we select the one coded using the smallest value
             grepl("1", edu_level) ~ "college or university degree",
             grepl("2", edu_level) ~ "A or AS levels",
             grepl("3", edu_level) ~ "O level or GCSEs",
             grepl("4", edu_level) ~ "CSEs",
             grepl("5", edu_level) ~ "NVQ or HND or HNC",
             grepl("6", edu_level) ~ "other professional qualifications"),
         edu_level = 
           factor(
             edu_level,
             levels = c("college or university degree", "A or AS levels", "O level or GCSEs", "CSEs", 
                        "NVQ or HND or HNC", "other professional qualifications", "none of the above", "prefer not to answer"),
             labels = c("college or university degree", "A or AS levels", "O level or GCSEs", "CSEs", 
                        "NVQ or HND or HNC", "other professional qualifications", "none of the above", "prefer not to answer")
           )
         )

baseline %>%
  janitor::tabyl(edu_level)

# collapsed version of edu_level
# people who did not answer or responded "prefer not to answer" will be coded missing 
baseline %<>%
  mutate(edu_level_na = 
           fct_recode(
             edu_level,
             "college or university degree" = "college or university degree",
             "less than college" = "A or AS levels", 
             "less than college" = "O level or GCSEs",
             "less than college" = "CSEs",
             "professional degree" = "NVQ or HND or HNC",
             "professional degree" = "other professional qualifications",
             "none of the above" = "none of the above"
           ),
         edu_level_na = 
           factor(
             edu_level_na,
             levels = c("less than college", "college or university degree", "professional degree", "none of the above"),
             labels = c("less than college", "college or university degree", "professional degree", "none of the above")
           ),
         edu_level_col = 
           fct_explicit_na(
             edu_level_na,
             na_level = "missing")
         ) %>%
  dplyr::select(id, age, yob, mob, sex, cur_emp, cur_emp_na, cur_emp_col, edu_level, edu_level_na, edu_level_col, everything())

table(baseline$edu_level_na, useNA = "ifany")
table(baseline$edu_level_col, useNA = "ifany")
```

## Average household income    
```{r clean_01 income}
table(baseline$`738-0.0`, useNA = "ifany")

baseline %<>%
  rename(income = `738-0.0`) %>%
  mutate(income_na = 
           case_when(
             income == "-1" ~ "",
             income == "-3" ~ "",
             income == "1" ~ "less than 18000",
             income == "2" ~ "18000 to 30999",
             income == "3" ~ "31000 to 51999",
             income == "4" ~ "52000 to 100000", 
             income == "5" ~ "greater than 100000"),
         income_na = 
           factor(
             income_na,
             levels = c("less than 18000", "18000 to 30999", "31000 to 51999", "52000 to 100000", "greater than 100000"),
             labels = c("less than 18000", "18000 to 30999", "31000 to 51999", "52000 to 100000", "greater than 100000")
           ),
         income = 
           fct_explicit_na(
             income_na,
             na_level = "missing")
         )

table(baseline$income_na, useNA = "ifany")
table(baseline$income, useNA = "ifany")
```

## Physical activity  
### IPAQ activity group  
```{r clean_01 ipaq}
table(baseline$`22032-0.0`, useNA = "ifany")

baseline %<>%
  rename(ipaq_cat = `22032-0.0`) %>%
  mutate(ipaq_cat_na =
           factor(
             ipaq_cat,
             levels = c(0:2),
             labels = c("low", "moderate", "high")),
         ipaq_cat = 
           fct_explicit_na(
             ipaq_cat_na,
             na_level = "missing")) %>%
  dplyr::select(id, age, yob, mob, sex, cur_emp, cur_emp_na, cur_emp_col, edu_level, edu_level_na, edu_level_col, ipaq_cat_na, ipaq_cat,
                everything())

table(baseline$ipaq_cat_na, useNA = "ifany")
table(baseline$ipaq_cat, useNA = "ifany")
```

### Above moderate/vigorous recommendation
```{r clean_01 pa moderate to vigorous}
table(baseline$`22035-0.0`, useNA = "ifany")

baseline %<>%
  rename(pa_mod_vig = `22035-0.0`) %>%
  mutate(pa_mod_vig_na =
           factor(
             pa_mod_vig,
             levels = c(0, 1), 
             labels = c("no", "yes")
           ),
         pa_mod_vig = 
           fct_explicit_na(
             pa_mod_vig_na,
             na_level = "missing"
           )) %>%
  dplyr::select(id, age, yob, mob, sex, cur_emp, cur_emp_na, cur_emp_col, edu_level, edu_level_na, edu_level_col, ipaq_cat_na, ipaq_cat,  
                pa_mod_vig_na, pa_mod_vig, everything())

table(baseline$pa_mod_vig_na, useNA = "ifany")
table(baseline$pa_mod_vig, useNA = "ifany")
```

### Above moderate/vigorous/walking recommendations  
```{r clean_01 pa walking moderate vigorous}
table(baseline$`22036-0.0`, useNA = "ifany")

baseline %<>%
  rename(pa_mod_vig_walk = `22036-0.0`) %>%
  mutate(pa_mod_vig_walk_na =
           factor(
             pa_mod_vig_walk,
             levels = c(0, 1), 
             labels = c("no", "yes")
           ),
         pa_mod_vig_walk = 
           fct_explicit_na(
             pa_mod_vig_walk_na,
             na_level = "missing"
           )) %>%
  dplyr::select(id, age, yob, mob, sex, cur_emp, cur_emp_na, cur_emp_col, edu_level, edu_level_na, edu_level_col, ipaq_cat_na, ipaq_cat,  
                pa_mod_vig_na, pa_mod_vig, pa_mod_vig_walk_na, pa_mod_vig_walk, everything())

table(baseline$pa_mod_vig_walk_na, useNA = "ifany")
table(baseline$pa_mod_vig_walk, useNA = "ifany")
```

### MET minutes per week for walking, for moderate, vigorous activity, and for all activity     
```{r clean_01 met minutes}
baseline %<>%
  rename(
    met_min_walk = `22037-0.0`,
    met_min_mod = `22038-0.0`, 
    met_min_vig = `22039-0.0`,
    met_min_total = `22040-0.0`
    )
```

### Summed days and minutes of activity  
```{r clean_01 pa summed}
baseline %<>%
  rename(
    pa_days = `22033-0.0`,
    pa_minutes = `22034-0.0`
    )
```

## Smoking  
### Ever smoked  
```{r clean_01 ever smoked}
table(baseline$`20160-0.0`, useNA = "ifany")

baseline %<>%
  rename(smk_ever = `20160-0.0`) %>%
  mutate(smk_ever_na =
           factor(
             smk_ever,
             levels = c(0, 1), 
             labels = c("no", "yes")
           ),
         smk_ever = 
           fct_explicit_na(
             smk_ever_na,
             na_level = "missing"
           )) %>%
  dplyr::select(id, age, yob, mob, sex, cur_emp, cur_emp_na, cur_emp_col, edu_level, edu_level_na, edu_level_col, ipaq_cat_na, ipaq_cat,  
                pa_mod_vig_na, pa_mod_vig, pa_mod_vig_walk_na, pa_mod_vig_walk, met_min_mod, met_min_vig, met_min_walk, met_min_total, 
                pa_days, pa_minutes, smk_ever_na, smk_ever, everything())

table(baseline$smk_ever_na, useNA = "ifany")
table(baseline$smk_ever, useNA = "ifany")
```

### Pack years of smoking  
```{r clean_01 smk pack-years}
baseline %<>%
  rename(smk_py = `20161-0.0`)
```

### Pack years of adult smoking as proportion of life span exposed to smoking  
```{r clean_01 smk proportion}
baseline %<>%
  rename(smk_prop = `20162-0.0`)
```

### Smoking status  
```{r clean_01 smk}
table(baseline$`20116-0.0`, useNA = "ifany")

baseline %<>%
  rename(smk = `20116-0.0`) %>%
  mutate(smk_na = 
           factor(
             smk,
             levels = c(0:2),
             labels = c("never", "former", "current")
           ),
         smk = 
           fct_explicit_na(
             smk_na,
             na_level = "missing"
           ))

table(baseline$smk_na, useNA = "ifany")
table(baseline$smk, useNA = "ifany")
```

## Alcohol intake  
### Alcohol drinker status  
```{r clean_01 alc}
table(baseline$`20117-0.0`, useNA = "ifany")

baseline %<>%
  rename(alc = `20117-0.0`) %>%
  mutate(alc_na = 
           factor(
             alc,
             levels = c(0:2),
             labels = c("never", "former", "current")
           ),
         alc = 
           fct_explicit_na(
             alc_na,
             na_level = "missing"
           ))

table(baseline$alc_na, useNA = "ifany")
table(baseline$alc, useNA = "ifany")
```

### Alcohol intake frequency  
```{r clean_01 alc freq}
table(baseline$`1558-0.0`, useNA = "ifany")

baseline %<>%
  rename(alc_freq = `1558-0.0`) %>%
  mutate(alc_freq_na = 
           factor(
             alc_freq,
             levels = c(6:1),
             labels = c("never", "special occasions only", "one to three times a month", 
                        "once or twice a week", "three or four times a week", "daily or almost daily")
           ),
         alc_freq = 
           fct_explicit_na(
             alc_freq_na,
             na_level = "missing"
           ))

table(baseline$alc_freq_na, useNA = "ifany")
table(baseline$alc_freq, useNA = "ifany")
```

## Medication  
* `6177-0.0` field is for men and `6153-0.0` field is for women  
* The responses from both fields were combined to create each variable for statin, bp medication, insulin, HRT, and OC separately.  

```{r clean_01 med}
table(baseline$`6177-0.0`, useNA = "ifany")
table(baseline$`6153-0.0`, useNA = "ifany")
```

### Statin  
```{r clean_01 statin}
baseline %<>%
  mutate(statin = 
           case_when(
             `6177-0.0` == "-1" | `6153-0.0` == "-1" ~ "",
             `6177-0.0` == "-3" | `6153-0.0` == "-3" ~ "",
             `6177-0.0` == "-7" | `6153-0.0` == "-7" ~ "0",
             grepl("1", `6177-0.0`) ~ "1", 
             grepl("1", `6153-0.0`) ~ "1",
             TRUE ~ "0"
           ), 
         statin = as.numeric(statin),
         statin_na = 
           factor(
             statin,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         statin = 
           fct_explicit_na(
             statin_na,
             na_level = "missing"
           ))

table(baseline$statin_na, useNA = "ifany")
table(baseline$statin, useNA = "ifany")

baseline %>%
  janitor::tabyl(statin, sex)
```

### Blood pressure medication  
```{r clean_01 bp med}
baseline %<>%
  mutate(bp_med = 
           case_when(
             `6177-0.0` == "-1" | `6153-0.0` == "-1" ~ "",
             `6177-0.0` == "-3" | `6153-0.0` == "-3" ~ "",
             `6177-0.0` == "-7" | `6153-0.0` == "-7" ~ "0",
             grepl("2", `6177-0.0`) ~ "1", 
             grepl("2", `6153-0.0`) ~ "1",
             TRUE ~ "0"
           ), 
         bp_med = as.numeric(bp_med), 
         bp_med_na = 
           factor(
             bp_med,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         bp_med = 
           fct_explicit_na(
             bp_med_na,
             na_level = "missing"
           ))

table(baseline$bp_med_na, useNA = "ifany")
table(baseline$bp_med, useNA = "ifany")

baseline %>%
  janitor::tabyl(bp_med, sex)
```

### Insulin  
```{r clean_01 insulin}
baseline %<>%
  mutate(insulin = 
           case_when(
             `6177-0.0` == "-1" | `6153-0.0` == "-1" ~ "",
             `6177-0.0` == "-3" | `6153-0.0` == "-3" ~ "",
             `6177-0.0` == "-7" | `6153-0.0` == "-7" ~ "0",
             grepl("3", `6177-0.0`) ~ "1", 
             grepl("3", `6153-0.0`) ~ "1",
             TRUE ~ "0"
           ), 
         insulin = as.numeric(insulin),
         insulin_na = 
           factor(
             insulin,
             levels = c("0", "1"),
             labels = c("no", "yes")
           ),
         insulin = 
           fct_explicit_na(
             insulin_na,
             na_level = "missing"
           ))

table(baseline$insulin_na, useNA = "ifany")
table(baseline$insulin, useNA = "ifany")

baseline %>%
  janitor::tabyl(insulin, sex)
```

### Hormone replacement therapy    
```{r clean_01 hrt}
baseline %<>%
  mutate(hrt = 
           case_when(
             sex == "men" ~ "",
             `6153-0.0` == "-1" ~ "",
             `6153-0.0` == "-3" ~ "",
             `6153-0.0` == "-7" ~ "0",
             grepl("4", `6153-0.0`) ~ "1", 
             TRUE ~ "0"
           ), 
         hrt = as.numeric(hrt),
         hrt_na = 
           factor(
             hrt,
             levels = c("0", "1"),
             labels = c("no", "yes")
           ),
         hrt = 
           fct_explicit_na(
             hrt_na,
             na_level = "missing"
           ))

table(baseline$hrt_na, useNA = "ifany")
table(baseline$hrt, useNA = "ifany")

baseline %>%
  janitor::tabyl(hrt, sex)
```

### Oral contraceptive pill or minipill  
```{r clean_01 oc}
baseline %<>%
  mutate(oc = 
           case_when(
             sex == "men" ~ "",
             `6153-0.0` == "-1" ~ "",
             `6153-0.0` == "-3" ~ "",
             `6153-0.0` == "-7" ~ "0",
             grepl("5", `6153-0.0`) ~ "1", 
             TRUE ~ "0"
           ), 
         oc = as.numeric(oc),
         oc_na = 
           factor(
             oc,
             levels = c("0", "1"),
             labels = c("no", "yes")
           ),
         oc = 
           fct_explicit_na(
             oc_na,
             na_level = "missing"
           ))

table(baseline$oc_na, useNA = "ifany")
table(baseline$oc, useNA = "ifany")

baseline %>%
  janitor::tabyl(oc, sex)
```

## Blood pressure  
```{r clean_01 bp}
baseline %<>%
  rename(
    dbp_1 = `4079-0.0`,
    dbp_2 = `4079-0.1`,
    sbp_1 = `4080-0.0`,
    sbp_2 = `4080-0.1`
    )
```

## Physical measures
### Height  
```{r clean_01 height}
baseline %<>%
  rename(height = `50-0.0`)
```

### Weight  
```{r clean_01 weight}
baseline %<>%
  rename(weight = `21002-0.0`)
```

### Body mass index  
```{r clean_01 bmi}
baseline %<>%
  rename(bmi = `21001-0.0`)
```

## Blood cell counts and percentages  
### White blood cells and diff counts  
```{r clean_01 wbc diff}
baseline %<>%
  rename(
    wbc = `30000-0.0`,
    neutro = `30140-0.0`,
    neutro_pct = `30200-0.0`,
    lymph = `30120-0.0`, 
    lymph_pct = `30180-0.0`,
    mono = `30130-0.0`,
    mono_pct = `30190-0.0`,
    baso = `30160-0.0`, 
    baso_pct = `30220-0.0`,
    eos = `30150-0.0`, 
    eos_pct = `30210-0.0`
    ) %>%
  mutate(
    ln_wbc = log(wbc + 1),
    ln_neutro = log(neutro + 1),
    ln_lymph = log(lymph + 1),
    ln_mono = log(mono + 1),
    ln_baso = log(baso + 1),
    ln_eos = log(eos + 1)
    )
```

### Red blood cells  
```{r clean_01 rbc}
baseline %<>%
  rename(rbc = `30010-0.0`,
         rbcdw = `30070-0.0`,
         n_rbc = `30170-0.0`,
         n_rbc_pct = `30230-0.0`) %>%
  mutate(ln_rbc = log(rbc + 1))
```

### Hematocrit, Hb  
```{r clean_01 rbc features}
baseline %<>%
  rename(hb = `30020-0.0`, 
         hct = `30030-0.0`, 
         mcv = `30040-0.0`,
         mch = `30050-0.0`,
         mchc = `30060-0.0`)
```

### Reticulocytes  
```{r clean_01 reti}
baseline %<>%
  rename(reti = `30250-0.0`,
         reti_pct = `30240-0.0`,
         hls_reti = `30300-0.0`, 
         hls_reti_pct = `30290-0.0`,
         reti_immature = `30280-0.0`,
         mrv = `30260-0.0`,
         mscv = `30270-0.0`)
```

### Platelets  
```{r clean_01 plt}
baseline %<>%
  rename(plt = `30080-0.0`,
         platcrit = `30090-0.0`,
         mpv = `30100-0.0`,
         pdw = `30110-0.0`) %>%
  mutate(ln_plt = log(plt + 1))
```

## Blood chemistry  
### Liver function tests  
```{r clean_01 lft}
baseline %<>%
  rename(albumin = `30600-0.0`, 
         alp = `30610-0.0`,
         alt = `30620-0.0`,
         ast = `30650-0.0`, 
         dbil = `30660-0.0`,
         ggt = `30730-0.0`,
         tbil = `30840-0.0`,
         tprotein = `30860-0.0`)
```

### Kidney function tests  
```{r clean_01 kft}
baseline %<>%
  rename(cr = `30700-0.0`,
         cystatin_c = `30720-0.0`,
         urea = `30670-0.0`)
```

### Glucose and HbA1c  
```{r clean_01 gluc}
baseline %<>%
  rename(glucose = `30740-0.0`,
         hba1c = `30750-0.0`)
```

### Lipids and lipoprotiens  
```{r clean_01 lipids}
baseline %<>%
  rename(apoa = `30630-0.0`, 
         apob = `30640-0.0`,
         cholesterol = `30690-0.0`,
         hdl = `30760-0.0`,
         ldl_direct = `30780-0.0`,
         lipoprotein_a = `30790-0.0`,
         trig = `30870-0.0`)
```

### Others   
```{r clean_01 others} 
baseline %<>%
  rename(calcium = `30680-0.0`,
         igf1 = `30770-0.0`,
         estradiol = `30800-0.0`,
         phosphate = `30810-0.0`,
         rf = `30820-0.0`,
         shbg = `30830-0.0`,
         testosterone = `30850-0.0`, 
         urate = `30880-0.0`,
         vitd = `30890-0.0`,
         crp = `30710-0.0`)
```

## Infectious disease labs  
### Ebstein-Barr virus  
```{r clean_01 ebv}
table(baseline$`23053-0.0`, useNA = "ifany")

baseline %<>%
  rename(ebv = `23053-0.0`) %>%
  mutate(ebv =
           factor(
             ebv,
             levels = c(0, 1), 
             labels = c("no", "yes")
           ),
         ebv = 
           fct_explicit_na(
             ebv,
             na_level = "missing"
           )) 

baseline %>%
  janitor::tabyl(ebv)
```

### CMV  
```{r clean_01 cmv}
table(baseline$`23054-0.0`, useNA = "ifany")

baseline %<>%
  rename(cmv = `23054-0.0`) %>%
  mutate(cmv =
           factor(
             cmv,
             levels = c(0, 1), 
             labels = c("no", "yes")
           ),
         cmv = 
           fct_explicit_na(
             cmv,
             na_level = "missing"
           )) 

baseline %>%
  janitor::tabyl(cmv)
```

## Self-reported health outcomes  
```{r clean_01 dx_self}
table(baseline$`6150-0.0`, useNA = "ifany")
```

### Heart attack diagnosed by doctor  
```{r clean_01 dx_self_heart_attack}
baseline %<>%
  mutate(dx_self_heart_attack = 
           case_when(
             grepl("1", `6150-0.0`) ~ "1", 
             grepl("-3", `6150-0.0`) ~ "", 
             `6150-0.0` == "" ~ "",
             TRUE ~ "0"
           ), 
         dx_self_heart_attack =
           as.numeric(dx_self_heart_attack), 
         dx_self_heart_attack = 
           factor(
             dx_self_heart_attack,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         dx_self_heart_attack = 
           fct_explicit_na(
             dx_self_heart_attack,
             na_level = "missing"
           ))

baseline %>%
  janitor::tabyl(dx_self_heart_attack)

baseline %<>%
  rename(dx_self_age_heart_attack = `3894-0.0`)
```

### Angina  
```{r clean_01 dx_self_angina}
baseline %<>%
  mutate(dx_self_angina = 
           case_when(
             grepl("2", `6150-0.0`) ~ "1", 
             grepl("-3", `6150-0.0`) ~ "",
             `6150-0.0` == "" ~ "", 
             TRUE ~ "0"
           ), 
         dx_self_angina =
           as.numeric(dx_self_angina),
         dx_self_angina = 
           factor(
             dx_self_angina,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         dx_self_angina = 
           fct_explicit_na(
             dx_self_angina,
             na_level = "missing"
           ))

baseline %>%
  janitor::tabyl(dx_self_angina)

baseline %<>%
  rename(dx_self_age_angina = `3627-0.0`)
```

### Stroke  
```{r clean_01 dx_self_stroke}
baseline %<>%
  mutate(dx_self_stroke = 
           case_when(
             # the order of grepl is different here because "3" run before "-3" ends up recoding "-3" as dx_self_stroke == 1
             grepl("-3", `6150-0.0`) ~ "",
             grepl("3", `6150-0.0`) ~ "1", 
             `6150-0.0` == "" ~ "", 
             TRUE ~ "0"
           ), 
         dx_self_stroke =
           as.numeric(dx_self_stroke),
         dx_self_stroke = 
           factor(
             dx_self_stroke,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         dx_self_stroke = 
           fct_explicit_na(
             dx_self_stroke,
             na_level = "missing"
           ))

baseline %>%
  janitor::tabyl(dx_self_stroke)

baseline %<>%
  rename(dx_self_age_stroke = `4056-0.0`)
```

### High blood pressure  
```{r clean_01 self_dx_htn}
baseline %<>%
  mutate(dx_self_htn = 
           case_when(
             grepl("4", `6150-0.0`) ~ "1", 
             grepl("-3", `6150-0.0`) ~ "",
             `6150-0.0` == "" ~ "",
             TRUE ~ "0"
           ), 
         dx_self_htn =
           as.numeric(dx_self_htn),
         dx_self_htn = 
           factor(
             dx_self_htn,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         dx_self_htn = 
           fct_explicit_na(
             dx_self_htn,
             na_level = "missing"
           ))

baseline %>%
  janitor::tabyl(dx_self_htn)

baseline %<>%
  rename(dx_self_age_htn = `2966-0.0`)
```

### Diabetes  
```{r clean_01 self_dx_diabetes}
table(baseline$`2443-0.0`, useNA = "ifany")

baseline %<>%
  mutate(dx_self_diabetes = 
           case_when(
             `2443-0.0` == -1 | `2443-0.0` == -3 ~ "", 
             grepl("1", `2443-0.0`) ~ "1", 
             grepl("0", `2443-0.0`) ~ "0",
             TRUE ~ ""
           ), 
         dx_self_diabetes =
           as.numeric(dx_self_diabetes),
         dx_self_diabetes = 
           factor(
             dx_self_diabetes,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         dx_self_diabetes = 
           fct_explicit_na(
             dx_self_diabetes,
             na_level = "missing"
           ))

baseline %>%
  janitor::tabyl(dx_self_diabetes)

baseline %<>%
  rename(dx_self_age_diabetes = `2976-0.0`)
```

### Cancer  
```{r clean_01 dx_self_cancer}
table(baseline$`2453-0.0`, useNA = "ifany")

baseline %<>%
  mutate(dx_self_cancer = 
           case_when(
             `2453-0.0` == -1 | `2453-0.0` == -3 ~ "", 
             grepl("1", `2453-0.0`) ~ "1", 
             grepl("0", `2453-0.0`) ~ "0",
             TRUE ~ ""
           ), 
         dx_self_cancer =
           as.numeric(dx_self_cancer),
         dx_self_cancer = 
           factor(
             dx_self_cancer,
             levels = c(0, 1),
             labels = c("no", "yes")
           ),
         dx_self_cancer = 
           fct_explicit_na(
             dx_self_cancer,
             na_level = "missing"
           ))

baseline %>%
  janitor::tabyl(dx_self_cancer)
```

## Remove unused variables  
```{r clean_01 remove variables}
baseline %<>%
  dplyr::select(-c(`137-0.0`, `845-0.0`, 
                   `1239-0.0`, `1249-0.0`, `1259-0.0`, `1269-0.0`, `1279-0.0`, `2443-0.0`, `2453-0.0`, `2463-0.0`, `2473-0.0`, 
                   `2644-0.0`, `2986-0.0`, `3005-0.0`, `3166-0.1`, `3436-0.0`, `3446-0.0`, `3731-0.0`,
                   `3761-0.0`, `3786-0.0`, `3992-0.0`, `4012-0.0`, `4022-0.0`, `4041-0.0`, 
                   `5959-0.0`, `6150-0.0`, `6151-0.0`, `6152-0.0`, `6153-0.0`, `6154-0.0`, `6155-0.0`, `6177-0.0`, `6179-0.0`, 
                   `12144-2.0`, `20003-0.0`, `90192-0.0`))
```

# Generate new variables  
## Indicator variable for outliers of blood cell counts - older version    
```{r clean_01 ct_outlier}
baseline %<>%  
  mutate(CT_Outlier = 
           case_when(
             log(wbc + 1) < 1 | log(wbc + 1) > 3 | 
               rbc < 1.9 | rbc > 7 | log(rbc + 1) < 1.2 | log(n_rbc + 1) > 0.75 | log(reti + 1) > 0.6 | 
               log(mrv + 1) < 4.0 | log(mrv + 1) > 5.25 | hb < 6 | hb > 20 | hct < 20 | hct > 60 | mcv < 50 | mcv > 125 | 
               log(mch + 1) < 2.9 | log(mch + 1) > 4 | log(mchc + 1) < 3.3 | log(mchc + 1) > 4 | log(rbcdw + 1) > 3.4 | 
               log(plt + 1) < 3 | log(plt + 1) > 7 | plt > 750 | log(platcrit + 1) > 0.5 | log(pdw + 1) < 2.75 | 
               log(mono + 1) > 2 | log(eos + 1) > 1.25 | log(baso + 1) > 0.75 ~ 1),
         CT_Outlier = ifelse(is.na(CT_Outlier), 0, CT_Outlier), 
         n_rbc_yn = ifelse(n_rbc == 0, 0, 1)) 
table(baseline$CT_Outlier)
```

## Indicator variable for outliers of blood cell counts - newer version  
* Same as the `ExcludeCells` Dan uses  

```{r clean_01 ct_exclude}
baseline %<>%  
  mutate(ct_exclude = 
           case_when(
             log(wbc + 1) > 1.25 & log(wbc + 1) < 3 & 
               log(rbc + 1) > 1.4 & log(rbc + 1) < 2 & 
               plt > 10 & plt < 500 & 
               log(lymph + 1) > 0.10 & log(lymph + 1) < 2 &
               log(mono + 1) < 0.9 &
               log(neutro + 1) > 0.75 & log(neutro) < 2.75 & 
               log(eos + 1) < 0.75 & log(baso + 1) < 0.45 ~ 0),
         ct_exclude = ifelse(is.na(ct_exclude), 1, ct_exclude)) 
table(baseline$ct_exclude)
```

## Convert units  
```{r clean_01 units}
baseline %<>%
  ## convert from nmol/L to mg/dL
  mutate(cholesterol_tr = cholesterol * (1/0.0259),
         glucose_tr     = glucose * (18.0182),
         hdl_tr         = hdl * (1/0.0259),
         ldl_tr         = ldl_direct * (1/0.0259),
         trig_tr        = trig * (1/0.0113)) %>%
  ## convert from mmol/mol to %
  mutate(hba1c_tr = (hba1c/10.929) + 2.15) 
```

## Calculated LDL  
```{r clean_01 ldl calculated}
baseline %<>%
  mutate(
    ldl_cal = ifelse(trig_tr < 400, cholesterol_tr - hdl_tr - (trig_tr/5), NA)
    )
```

# Save data set 
```{r clean_01 save}
baseline %>%
  skimr::skim_without_charts()

baseline %>%
  saveRDS("ukb_heteroplasmy_mortality_01_baseline.rds")
```
